### 2017-07-18 etc/uWUE_gw_notes/wue_gs.tex
added medlyn model, not need to test if scipy optmiziers can solve efficiently

\input{def}
%\usepackage{amsmath}
\usepackage{graphics, graphicx}
\graphicspath{ {./} } %directory for figures 
%\pagestyle{empty}

\begin{document}
\begin{center}
\large{Using uWUE to remove dependence of ET on GPP}\end{center}

\bigskip

Penman montieth is given by:

\begin{equation}
  \lambda E = \frac{\Delta A + \frac{\rho_a c_p D_{ref}}{r_a}}{\Delta + \gamma(1 + \frac{r_s}{r_a})}
\end{equation}

where:
\begin{equation}
  r_s = \frac{1}{LAI \cdot  g_{st}}
\end{equation}

Using the Medlyn model (from Frank et. al. 2017) we have:

\begin{equation}
g_{st} = g_{w} = 1.6 \left( 1 + \frac{g_{1M}}{\sqrt{D_s}} \right) \frac{A}{c_s}
\end{equation}

Here, $g_{1M}$ has units of kPa$^{0.5}$, $A$ is the net CO$_2$ assimilation rate in $\mu$ mol m$^{-2}$ s$^{-1}$, $c_s$ is the CO$_2$ concentration in ppm (supposed to be at leaf surface). The problem is the $A$ term that we need to get rid of. Also $g_w$ as above has units of mol (air?) m$^{-2}$ s$^{-1}$.

we can use:

\begin{equation}
  \text{uWUE} = \frac{\text{GPP} \cdot \sqrt{\text{VPD}}}{ET}
\end{equation}

This has units of $\frac{\text{g C  hPA}^{0.5}}{\text{kg H$_2$O}}$.

So let's do these unit conversions, starting with $g_w$ unit converstion. Note that the mean molecular mass of dry air is $28.9645$ g/mol. So then we have:

\begin{equation}
  g_{st\; -\; m/s} = g_w \frac{mol}{m^2 s^1} \frac{28.9645 g air}{ mol air} \frac{1 kg}{1000 g} \frac{m^3}{\rho \; kg} = \frac{m}{s}
\end{equation}

so:
\begin{equation}
  g_{st\; -\; m/s} = g_w \frac{R_{air}}{\rho_{air}}
\end{equation}

where:
\begin{equation}
  R_{air} = .0289645 \frac{\text{kg}}{\text{mol}}
\end{equation}

Now we just need to convert uWUE from gC to $\mu$ mol C:

\begin{equation}
uWUE_{\mu \; \text{mol C}} = uWUE \frac{ \text{mol C}}{12.011 \text{g C}} \frac{1.e6 \; \mu \; \text{mol}}{mol}
\end{equation}

so:
\begin{equation}
  uWUE = \frac{uWUE \;1.e6}{ 12.011}
\end{equation}

\end{document}