### 2017-08-29 etc/2017-08-28_VPD_uWUE_notes/explicit.tex
NOTE DELETING OLD PLOTTING CODE THAT COULD BE USEFUL, also found order of mangitude error in new model but still not 1:1

\input{def}
%\usepackage{amsmath}
\usepackage{graphics, graphicx}
\graphicspath{ {./} } %directory for figures 
%\pagestyle{empty}

\begin{document}
\begin{center}
\large{Using uWUE to remove dependence of ET on GPP}\end{center}

\bigskip

This document is to analytically caluclate ET and $\frac{\partial ET} {\partial VPD}$.

Penman montieth is given by:

\begin{equation}
  \lambda E = \frac{\Delta R + g_a \rho_a c_p D_{s}}{\Delta + \gamma(1 + \frac{g_a}{g_s})}
\end{equation}

where from Medlyn we have:
\begin{equation}
  g_s = \text{LAI } \frac{R T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{D_{s}}}\right) \frac{A}{c_s}
\end{equation}

Here, $g_{1M}$ has units of kPa$^{0.5}$, $A$ is the net CO$_2$ assimilation rate in $\mu$ mol m$^{-2}$ s$^{-1}$, $c_s$ is the CO$_2$ concentration in ppm (supposed to be at leaf surface). The problem is the $A$ term that we need to get rid of. Also $g_w$ as above has units of mol (air?) m$^{-2}$ s$^{-1}$.

we can use:

\begin{equation}
  \text{uWUE} = \frac{\text{GPP} \cdot \sqrt{\text{VPD}}}{ET}
\end{equation}

As long as we put uWUE into $\mu mol$, we can use the uWUE provided by Zhou et. al. To do this just multiply be a factor $ \frac{1 mol C}{12.011 g C} \frac{1.e6 \mu mol}{1 mol}$. Then make sure that units of VPD match. If using Franks et al. 2017 and Zhou et al. then VPD within the medlyn model should be in kPa, and VPD cancelling uWUE should be in hPa.

Then we can show that if we multiply $g_s$ by $\frac{ uWUE \; ET}{A \sqrt{D_s} L_v}$ (which is equivalent to 1 with unit conversion above), we can rearrange Penman Moneith to get:

\begin{equation}
  \lambda E = \frac{\Delta R + g_a \left( \rho_a c_p D_{s} - \frac{\gamma c_s \sqrt{D_s} L_v P }{\text{LAI } R \; T \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})} \right) }{ \Delta + \gamma}
\end{equation}

If we want to solve for LAI as a function of all other knowns, we have:
\begin{equation}
LAI  = \frac{g_a \gamma c_s \sqrt{D_s} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p D_{s}\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{D_s}})}
\end{equation}

Now taking the derivative w.r.t. $D_s$ we get:

\begin{equation}
\frac{\partial \;  \lambda E}{\partial \; D_s} = \frac{g_a}{\Delta + \gamma}   \left( \rho_a c_p - \frac{\gamma c_s L_v  P }{\text{LAI }1.6 \; R\; T \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{D_s}}{2 (g_1 + \sqrt{D_s})^2}\right) \right)
\end{equation}

\end{document}