### 2017-10-19 doc/glab_meeting_2017-10-20/vpd.tex
added glab presentation

\documentclass{beamer}
%
% Choose how your presentation looks.
%
% For more themes, color themes and font themes, see:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
%
\mode<presentation>
{
  \usetheme{default}      % or try Darmstadt, Madrid, Warsaw, ...
  \usecolortheme{default} % or try albatross, beaver, crane, ...
  \usefonttheme{default}  % or try serif, structurebold, ...
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
}

\usepackage[english]{babel}
 % \usepackage[utf8x]{inputenc}
\usepackage[style=authoryear]{biblatex}
\addbibresource{references.bib}

\title[Your Short Title]{Bayesian Neural  Networks and Probalistic Programming}
\author{Adam Massmann and Laureline Josset}
\institute{Water Center NN Meetings}
\date{Week 5: Oct. 17th, 2017}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section{References}
\begin{frame}{References}
  \AtNextBibliography{\small}
  \printbibliography
\end{frame}

% \begin{figure}
%   \includegraphics[width=3in]{ranganath.png}
%   \caption{Reproduced from \cite{ranganath2014}.}
% \end{figure}

% \begin{table}
% \centering
% \begin{tabular}{l|r}
% Item & Quantity \\\hline
% Widgets & 42 \\
% Gadgets & 13
% \end{tabular}
% \caption{\label{tab:widgets}An example table.}
% \end{table}

\end{document}