### 2017-10-20 doc/glab_meeting_2017-10-20/vpd.tex
added vpd, calling it quits fo rthe night

\documentclass{beamer}
%
% Choose how your presentation looks.
%
% For more themes, color themes and font themes, see:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
%
\mode<presentation>
{
  \usetheme{default}      % or try Darmstadt, Madrid, Warsaw, ...
  \usecolortheme{default} % or try albatross, beaver, crane, ...
  \usefonttheme{default}  % or try serif, structurebold, ...
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
}

\usepackage[english]{babel}
 % \usepackage[utf8x]{inputenc}
\usepackage[style=authoryear]{biblatex}
\addbibresource{references.bib}

\title[Your Short Title]{When does VPD drive or reduce ET?}
\author{Adam with Pierre and Changjie}
\institute{G-Lab Meeting}
\date{October 20th, 2017}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{When does VPD drive or reduce ET?}
  \begin{itemize}
  \item \textbf{Hydrometeorologists} would say that an increase in VPD (increase in \textbf{atmospheric demand}) would drive an \textbf{increase in ET}.
  \item However, \textbf{plant physioligists} know that plants have evolved to use stomata to conserve and regulate water use. So \textbf{stomata closure} in response to increases in VPD may \textbf{decrease ET}.
  \end{itemize}
\end{frame}

\begin{frame}{The question is, which effect dominates with an increase in VPD: plant response (decrease in ET) or atmospheric demand (increase in ET)?}
  \begin{itemize}
  \item We hypothesize it will be a function of plant type and the environment:
    \begin{itemize}
    \item Plants that are evolved to conserve water will tend to reduce ET with increases in VPD.
    \item However the environment can overwhelm plant response: at very high VPD the atmospheric demand will dominate and plants will not be able to consernve water, no matter how much they have evolved to do so.
    \end{itemize}
  \end{itemize}
\end{frame}

\section{References}
\begin{frame}{References}
  \AtNextBibliography{\small}
  \printbibliography
\end{frame}

% \begin{figure}
%   \includegraphics[width=3in]{ranganath.png}
%   \caption{Reproduced from \cite{ranganath2014}.}
% \end{figure}

% \begin{table}
% \centering
% \begin{tabular}{l|r}
% Item & Quantity \\\hline
% Widgets & 42 \\
% Gadgets & 13
% \end{tabular}
% \caption{\label{tab:widgets}An example table.}
% \end{table}

\end{document}