### 2018-03-26 doc/paper/vpd_et_paper.tex
finished idealized sign and scaling sections

% 11/23/2015
\documentclass[draft,linenumbers]{agujournal}
% \draftfalse
\usepackage{makecell}
\usepackage{multirow} %AM for tables
\usepackage{booktabs}
\drafttrue

\journalname{Agricultural and Forest Meteorology: When does VPD drive or reduce ET?}

\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}

\begin{document}

\title{When does vapor pressure deficit drive or reduce evapotranspiration?}

\authors{Adam Massmann\affil{a}, Pierre Gentine\affil{a}, Changjie Lin\affil{a,b}}

\affiliation{a}{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
\affiliation{b}{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic Engineering, Tsinghua University, Beijing, CN 100084}

\correspondingauthor{Adam Massmann}{206-919-1364; akm2203@columbia.edu}

\begin{keypoints}
\item ecohydrology, evapotranspiration, aridity, vapor pressure deficit, Clausius-Clapeyron relation, plant physiology, stomatal response
\end{keypoints}

\begin{abstract}
  Increases to vapor pressure deficit (VPD) and atmospheric demand for water are expected with rising atmospheric CO$_2$. While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water. Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET. We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
  The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD. Plants that are evolved or bred to prioritize primary production over water conservation (e.g. crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing). However for forest, grass, and shrub plant types, ET more frequently decreases than increases with rising VPD. Therefore, we may expect reductions in ET due to CO$_2$-induced increases in VPD. This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.

\end{abstract}

\section{Introduction}

Vapor pressure deficit (VPD) is expected to rise over continents in the future due to the combination of increased temperature and, depending on region, decreased relative humidity \citep{Byrne_2013}. Increases in VPD increase the atmospheric demand for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.

The opposing effects of increased atmospheric demand and higher stomatal stress lead to two possible perspectives for how evapotranspiration (ET) responds to shifts in VPD. The first, a hydrometeorological perspective, is that higher VPD increases atmospheric demand for water from the land surface, and this drives an increase in evapotranspiration (ET). This perspective is particularly relevant because potential evapotranspiration (PET), which is used in many drought indices and hydrometeorological studies \citep[e.g.,][]{Heim_2002, Scheff_2015}, typically only quantifies changes in atmospheric demand and fails to account for ecosystem response \citep{Swann_2016}. In reality, plants' stomata have evolved to optimally regulate the exchange of water and carbon, and tend to partially close in response to increased atmospheric dryness \citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This leads to a plant physiology perspective, in which an increase in VPD, particularly in well-watered soil conditions, may actually correspond to a decrease in ET because of stomatal closure \citep[e.g.][]{Rigden_2017}.  In other words, the  question ``When does VPD drive or reduce ET?'' can be related to whether plant regulation or atmospheric demand dominates ET response.

The ET response to changes in VPD alters water partitioning between the soil and atmosphere. If plant response reduces ET with atmospheric drying then soil moisture will be better conserved. This would seem a sensible evolutionary strategy to cope with aridity. If stomata were fully passive \citep [similar to soil pores, e.g. ][]{Or_2013}, increased atmospheric aridity would strongly reduce soil moisture \citep{Berg_2017}. In turn, this would further increase aridity as low soil moisture levels increase the Bowen ratio, and cause increased temperature and atmospheric drying \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013, Gentine_2016, Berg_2016}. This however would not seem to be a sensible strategy for plants from an evolutionary standpoint.

We can use intuition about plant water conservation strategy to hypothesize about ET response to changes in VPD. Plants and ecosystems that evolved to conserve water, such as arid shrubs or savannah, should be more likely to reduce ET with increasing VPD, and plants that have evolved or have been engineered to care little about water, such as crops, will be more likely to increase ET with increasing VPD. Atmospheric conditions must matter as well. At the ecosystem scale, there are limits to plant water conservation strategies. As atmospheric demand for water (VPD) increases, ecosystems should begin to reach their water conservation limits and might not be able to entirely limit ET flux to the atmosphere. At this stage any further increase in VPD will most likely drive a (limited) increase in ET, because the increase in atmospheric demand for water overwhelms the limited plant response to conserve water.

The objective of the present manuscript is to use reasonable approximations as a tool to develop intuition for plant response to atmospheric drying and evaluate the VPD dependence of ET. This intuition will aid interpretation of observations, full complexity models, and facilitate the disentanglement of complex land-atmosphere feedbacks. In the past, similiar approaches were used to understand interactions between stomatal conductance, evapotranspiration and the environment \citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at the time researchers' understandng of the form of VPD's effect on plant physiology was limited, so they could not explore the sensitivity of ET to VPD, including VPD's effect on stomatal conductance and plant function.

Recent results have drastically improved our understanding of VPD's impact on physiology. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal conductance ($g_s$), including VPD response, by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, and extended use of this model to the ecosystem scale in \citet{Medlyn_2017}. Additionally, \citet{Zhou_2014} demonstrated that a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\; \sqrt{VPD}}{ET}\right)$ properly captures a constant relationship between GPP, ET, and VPD over a diurnal cycle. uWUE is also remarkably well conserved in the growing season across space and time, within a PFT \citep{Zhou_2015}. These results provide novell tools to examine how ET varies in response to VPD, while explicitly including VPD's effect on stoamtal conductance for the first time. In this manuscript, we leverage these tools to derive the theoretical response of ET to VPD using a Penman Monteith framework. Our theory is validated and tested at multiple eddy-covariance stations spanning various climates and plant functional types.

\section{Materials and Methods}
\subsection{Data}
\label{data}
We use both meteorological and eddy-covariance data from the FLUXNET2015 database (data available at \sloppy https://fluxnet.fluxdata.org/data/fluxnet2015-dataset/ \sloppy), including all sites with at least four years of data and observations of the variables described in the methods section (Section \ref{methods}). Seventy-three sites met these requirements, and were grouped into nine plant functional types (PFT) according to the International Geosphere-Biosphere Programme vegetation classification scheme \citep{Loveland_1999}: cropland (CRO), grass (GRA), deciduous broadleaf forest (DBF), evergreen broadlead forest (EBF), evergreen needleleaf forest (ENF), mixed forest (MF), closed shrub (CSH), savannah (SAV), and woody savannah (WSA) (site locations in Figure \ref{map_fig}).

\begin{figure}
\centering
\includegraphics[width=\textwidth]{./map.pdf}
\caption{Plant functional type and location of FLUXNET2015 sites used in this analysis.}
\label{map_fig}
 \end{figure}

The purpose of this study is to examine ecosystem response to atmospheric drying, focusing on the growing season. To accomplish this, we filter and quality control the data using a similar procedure as \cite{Zhou_2015}:
\begin{itemize}
\item Only measured or highest (``good'') quality gapfilled data, according to quality control flags, are used.
\item To isolate the growing season, we only use days in which the average Gross Primary Productivity (GPP) exceeds 10\% of the observed 95th percentile of GPP for a given site. GPP is calculated using the nighttime respiration partitioning method.
\item We remove days with rain and the day following to avoid issues with rain interception and sensor saturation at high relative humidity (\cite{MEDLYN_2011}).
\end{itemize}
Additionally, as in \citet{Lin_2018}, we restrict data to the daytime, which is identified when downwelling shortwave radiation is greater than 50 W m$^{-2}$ and sensible heat flux is greater than 5 W m$^{-2}$. To reduce the chance of sensor saturation at high relative humidity, we remove all time steps for which VPD is less than .01 kPa, and to reduce errors at low windspeeds we remove all periods with wind magnitudes less than 0.5 m s$^{-1}$. Timesteps with negative observed GPP or ET are also removed, and we aggregate half hourly data to hourly averages to reduce noise \citep{Lin_2018}. After these quality control procedures, 434,765 upscaled hourly observations remain.

\subsection{Methods}
\label{methods}
The Penman-Monteith equation \citep [hereafter PM,][]{Penman_1948, Monteith_1965} estimates ET as a function of observable atmospheric variables and surface conductances:
\begin{linenomath*}
  \begin{equation}
    \label{orig_pen}
    ET = \frac{\Delta R_{net} + g_a \rho_a c_p VPD}{\Delta + \gamma(1 + \frac{g_a}{g_s})},
  \end{equation}
\end{linenomath*}
where $\Delta$ is the change in saturation vapor pressure with temperature, given by Clausius-Clapeyron ($\frac{d \; e_s}{d \; T}$), $R_{net}$ is the net radiation minus ground heat flux, $g_a$ is aerodynamic conductance, $\rho_a$ is air density, $c_p$ is specific heat of air at constant pressure, $\gamma$ is the psychometric constant, and $g_s$ is the stomatal conductance (Table \ref{definitions}).

\begin{table}
  \caption{Definition of symbols and variables, with citation for how values are calculated, if applicable.}
  \label{definitions}
\centering
\small
\begin{tabular}{l c c c}
\hline
 Variable & Description & Units & Citation \\
\hline
$e_s$  & saturation vapor pressure & Pa  & - \\ 
  $T$  & temperature  & K & - \\
  $P$  & pressure & Pa  & - \\
$\Delta$  & $\frac{\partial e_s}{\partial T}$ & Pa K$^{-1}$ & - \\
$R_{net}$  & net radiation at land surface minus ground heat flux & W m$^{-2}$   & - \\
  $g_a$  & aerodynamic conductance & m s$^{-1}$  & \citet{Shuttleworth_2012} \\
  $\rho_a$  & air density & kg m$^{-3}$  & - \\
  $c_p$  & specific heat capacity of air at constant pressure & J K$^{-1}$ kg$^{-1}$ & - \\
  $VPD$  & vapor pressure deficit & Pa  & - \\
  $\gamma$  & psychometric constant & Pa K$^{-1}$   & - \\
  $g_s$  & stomatal conductance & m s$^{-1}$  & \citet{MEDLYN_2011} \\
  $g_{l-s}$  & leaf-scale stomatal conductance & mol m$^{-2}$ s$^{-1}$  & \citet{MEDLYN_2011} \\
  $R$ & universal gas constant & J mol$^{-1}$ K$^{-1}$ & - \\
  $R_{air}$ & gas constant of air & J  K$^{-1}$ kg$^{-1}$ & - \\
  $LAI$ & leaf area index & -& - \\
  $\sigma$ & uncertainty parameter & -& - \\
  $c_a$ & CO$_2$ concentration & $\mu$ mol CO$_2$ mol$^{-1}$ air& - \\
  $\lambda$ & marginal water cost of leaf carbon & mol H$_2$O mol$^{-1}$ CO$_2$ & - \\
  $\Gamma$ & CO$_2$ compensation point & - & - \\
  $\Gamma^*$ & CO$_2$ compensation point without dark respiration & - & - \\
\hline
\end{tabular}
\end{table}

\citet{MEDLYN_2011} developed a model for stomatal conductance ($g_s$) by combining an optimal photosynthesis theory \citep{Farquhar_1980, Katul_2010} with an empirical approach, which describes the dependence of $g_s$ to VPD. This resulted in the following model for leaf-scale stomatal conductance:

\begin{linenomath*}
  \begin{equation}
  g_{s-leaf} = g_0 + 1.6 \left(1 + \frac{g_{1-leaf}}{\sqrt{VPD}}\right) \frac{A}{c_a},
  \end{equation}
\end{linenomath*}
where $g_{1-leaf}$ is a leaf-scale ``slope'' parameter, A is the net CO$_2$ assimilation rate, and $c_a$ is the atmospheric CO$_2$ concentration at the leaf surface. \cite{MEDLYN_2011} relate the slope parameter ($g_{1-leaf}$) to physical parameters as:
\begin{linenomath*}
  \label{slope}
  \begin{equation}
  g_{1-leaf} = \sqrt{\frac{3 \, \Gamma^* \, \lambda}{1.6}},\footnote{Note this expression has units of of (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to Pa$^{1/2}$ using the ideal gas law.}
  \end{equation}
\end{linenomath*}

where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis (without dark respiration), and $\lambda$ is the marginal water cost of leaf carbon ($\frac{\partial \; \text{transpiration}}{\partial \; A}$). So, $g_{1-leaf}$ is a leaf-scale term reflecting the tradeoff of water for carbon uptake. The higher $g_{1-leaf}$, the more open the stomata and the more they release water in exchange for carbon.

The Medlyn model for stomatal conductance has been shown to behave very well across PFTs \citep[][]{Lin_2015}, and has been successfully adopted to ecosystem scale analysis in \citet{Medlyn_2017}. In units of m s$^{-1}$, the ecosystem scale stomatal conductance is:

\begin{linenomath*}
  \begin{equation}
    g_s = \frac{R \,T}{P} \; 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{GPP}{c_a},
      \label{medlyn}
  \end{equation}
\end{linenomath*}

where GPP is the ecoysystem scale gross primary production, and $g_1$ is an ecosystem scale analogue to $g_{1-leaf}$. We solve for $g_1$ following \citet{Medlyn_2017} (Equation 5), and take the median $g_1$ value to be representative of each PFT (Table \ref{pft}).

\begin{table}
  \caption{Plant functional types, their abbreviation, calculated Medlyn coefficient, and calculated uWUE. uWUE from \citet{Zhou_2015}, including the  observed standard deviation is shown for comparison. Note that uWUE from \citet{Zhou_2015} is calculated from a different set of sites, and that units are converted such that the quantities work with Equations 1-8 and the variables defined Table \ref{definitions}.}
    \small
\label{pft}
\centering
\begin{tabular}{l c c @{\qquad} c c}
  \hline
  \multirow{2}[3]{*}{Abbreviation} & \multirow{2}[3]{*}{PFT} & \multirow{2}[3]{*}{$g_1$ (Pa$^{0.5}$)} & \multicolumn{2}{c}{uWUE ($\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET])}  \\
  \cmidrule{4-5}

  & & & fitted & \citet{Zhou_2015} \\

  \hline
\input{pft_params}
  \hline
\end{tabular}
\end{table}

While Equation 3 can be used in PM (Equation 1), it will make analytical work with the function intractable because $GPP$ is functionally related to ET itself. Additionally, a perturbation to VPD should induce a physiological plant response that will alter GPP and cause an indirect change in stomatal conductance, in addition to the direct effect of VPD in Equation \ref{medlyn}. Therefore, in order to derive the response of ET to VPD, we must account for the functional relationship between GPP, ET, and VPD, and its effect on stomatal conductance. We can use aforementioned semi-empirical results of \citet{Zhou_2015} as a tool to approach this problem. \citet{Zhou_2015}, showed that the underlying Water Use Efficiency (uWUE):

\begin{linenomath*}
  \begin{equation}
    uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET}
    \label{uwue}
  \end{equation}
\end{linenomath*}

is relatively constant across time and moisture conditions within a plant functional type, and correctly captures a constant relationship between GPP, ET and VPD over a diurnal cycle \citep{Zhou_2014}. We can use uWUE to remove the $GPP$ dependence from $g_s$ in a way that makes the PM equation analytically tractable:

\begin{linenomath*}
  \begin{equation}
    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{\sqrt{VPD}}\right) \frac{uWUE \; ET}{c_a \; \sqrt{VPD}}.
    \label{new_g_s}
  \end{equation}
\end{linenomath*}

Plugging Equation \ref{new_g_s} into Equation \ref{orig_pen} and rearranging gives a new explicit expression for PM, in which dependence on $GPP$ is removed:

\begin{linenomath*}
  \begin{equation}
    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
    \label{et}
  \end{equation}
\end{linenomath*}

With Equation \ref{et} we have derived for the first time, using recent results \citep[][]{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017}, ET explicitly as function of environmental variables and two plant-specific constants, the slope parameter ($g_1$), and uWUE, both reflecting water conservation strategy. The slope parameter is related to the willingness of stomata to trade water for CO$_2$ and to keep stomata open. uWUE is a semi-empirical ecosystem-scale constant related to how WUE changes with VPD (specifically $VPD^{-1/2}$). It is also roughly proportional to physical constants:

\[uWUE \appropto \sqrt{\frac{c_a - \Gamma}{1.6 \lambda}},\]

where $\Gamma$ is the CO$_2$ compensation point \citep[Equation 5 in][]{Zhou_2014}. So uWUE is related to atmospheric CO$_2$ concentration and compensation point, and is inversely proportional to the marginal water cost of leaf carbon.

Given eddy-covariance FLUXNET2015 data (Section \ref{data}), every term in our new version of PM (Equation \ref{et}) is observed, except for uWUE, which we fit by calculating its expectation, given the model and FLUXNET2015 data.

However, eddy-covariance data are inherently noisy so we include a measure of uncertainty in our analysis. To account for observational error, as well as model uncertainty (e.g. temporal and spatial variations of uWUE and $g_1$), we introduce an uncertainty parameter $\sigma$ modifying uWUE:

\begin{linenomath*}
  \begin{equation}
    ET = \frac{\Delta R_{net} + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_a \sqrt{VPD} }{ R \; 1.6\; \sigma \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}
    \label{et_sigma}
  \end{equation}
\end{linenomath*}

Now, from each FLUXNET2015 observation (i.e. for each hourly observation at every time step) we can evaluate $\sigma$:

\begin{linenomath*}
  \begin{equation}
\sigma = - \frac{g_a \gamma c_a \sqrt{VPD} L_v P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R_{net} - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})},
    \label{sigma}
  \end{equation}
\end{linenomath*}

Thus, with this uncertainty analysis we can evaluate departure from our theory in observations, as a departure of $\sigma$ from unity. The variability of $\sigma$ across sites and time provides a measure of uncertainty in our model, assumptions, as well as in the FLUXNET2015 observations themselves. The variability of $\sigma$ then propagates through any uncertainty to our derivative of Equation \ref{et_sigma}:

\begin{linenomath*}
  \begin{equation}
    \frac{\partial \;  ET}{\partial \; VPD} = \frac{2\; g_a \; P}{T(\Delta + \gamma)}   \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_a }{1.6 \; R\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)
    \label{d_et}
  \end{equation}
\end{linenomath*}

With Equation \ref{d_et} we have an analytical framework for ecosystem response to atmospheric demand perturbations, with an implicit assumption of non-extreme drought conditions associated with our constant $uWUE$ assumption (e.g. we might expect $uWUE$ to decrease in extreme drought). There are a few subtleties to taking the derivative in Equation \ref{d_et}: $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while taking the derivative we evaluate $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET} {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD} \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH} \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$ fixed}}$. $RH$ and $e_s$ are assumed to be approximately independent, which is supported by the data (not shown).

We note one final comment on our derivation which is relevant for drought indices. If we approximate $c_a$ at a global mean CO$_2$ concentration, then the RHS of Equation \ref{et} is fully defined using commonly available weather station data and the constants published here. This makes Equation \ref{et} a useful alternative to PET in drought indices and hydrometeorological analysis for vegetated surfaces. Equation \ref{et} better reflects the physics of water exchange at the land surface and would only require fitting of uWUE and $g_1$, and can also account for changes in CO$_2$ concentration (assuming some relationship between $\lambda$ and [CO$_2$]), contrary to typical drought indices such as the Palmer Drought Severity Index (PDSI) \citep{Swann_2016}.

\section{Results}
\label{results}

By construction, the variability in the $\sigma$ term (Equation \ref{sigma}) contains all model and observational uncertainties. For an observation that perfectly matches our model and constant uWUE assumption, $\sigma$ will be one. Therefore, for our assumptions and framework to be reasonable $\sigma$ should be close to 1. An additional concern is that $\sigma$ may in fact be correlated with $VPD$, in which case the dependence would need to be accounted for when taking the derivative. Fortunately, there is a very weak dependence of $\sigma$ on VPD in their joint distrubution, and $\sigma$ is indeed close to unity i.e. $O(1)$ (Figure \ref{joint_vpd_sigma}). Given this weak dependence and the distribution of $\sigma$ we have confidence in our model framework and the data quality.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{./joint_vpd_sigma.pdf}
\caption{The joint distribution of $VPD$ and $\sigma$, with outliers removed (defined as lowest and highest 5\% of $\sigma$). $\sigma$ exhibits a weak dependence on $VPD$, and $\sigma$ is $O(1)$ for the bulk of the observations.}
\label{joint_vpd_sigma}
\end{figure}

Before calculating the sensitivity of ET to VPD, we will consider the functional form of Equation \ref{d_et}. There are two main terms: a ``scaling'' term, which modifies the magnitude but not the sign of the ET response to VPD  ($\frac{\partial \; ET}{\partial \; VPD}$):

\begin{equation}
  \frac{g_a \; P}{T(\Delta + \gamma)},
\end{equation}

and a ``sign'' term, which determines whether ET increases or decreases with VPD (i.e. atmospheric demand driven or physiologically controlled):

\begin{equation}
  \label{sign}
  \frac{c_p}{R_{air}} - \frac{\gamma c_a }{1.6 \; R\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right).
\end{equation}
All variables are positive, so the relative magnitude between the first term and the second term in the sign term (Equation \ref{sign}) will determine whether ET increases or decreases with increasing VPD. If the second term is larger then plant control dominates and ET decreases with increasing VPD. However, if the first term is larger, then atmospheric demand dominates and ET increases with increasing VPD.

\subsection{Functional Form of the Sign Term}
\label{sign_term}
First, we explore the variables within the sign term to gain better intuition on the driver of either the increase or reduction of ET with VPD. CO$_2$ concentration ($c_a$) and the psychometric constant ($\gamma$) are relatively constant over the dataset considered here so that the variability is dominated by $\sigma$ and $VPD$. uWUE could vary with soil moisture but has been shown to be relatively constant \citep{Zhou_2015}. This then means that the sign term only depends on VPD for a given PFT and is approximately just a function of $VPD$. We can further determine a critical threshold separating an increase from a decrease in ET, i.e. the threshold $VPD_{crit}$ such that the derivative vanishes $\frac{\partial \; ET}{\partial \; VPD} = 0$:
\small
\begin{linenomath*}
  \begin{equation}
VPD_{crit} = \frac{R_{air}}{4 c_p} \left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + \sqrt{\frac{\gamma c_a}{1.6\; R \;  uWUE}\left( \frac{\gamma c_a}{1.6\; R \;  uWUE} + 8 g_1 \frac{c_p}{R_{air}}\right)} - 4 g_1 \frac{c_p}{R_{air}} \right),
\label{vpd_min_et}
  \end{equation}
\end{linenomath*}
\normalsize
noting that $VPD_{crit}$ mostly depends on the PFT parameters uWUE and $g_1$, and only varies weakly with climate as most other parameters related to the environment are nearly constant. The calculated value of $VPD_{crit}$ for each PFT is shown in Table \ref{vpd_crit}. For any values of $VPD$ less than $VPD_{crit}$, ET will decrease with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} < 0$), and for values of $VPD$ greater than $VPD_{crit}$, ET will increase with increasing VPD ($\frac{\partial \; ET}{\partial \; VPD} > 0$). In other words, ecosystems regulate and mitigate evaporative losses up to the VPD limit, $VPD_{crit}$, above which atmospheric demand is just too high to be entirely compensated by stomatal and ecosystem regulation. We note however that even though ET increases again above the critical threshold, $VPD_{crit}$, ET is still much lower that potential evaporation as stomata are still strongly regulating vapor fluxes to the atmosphere. Indeed, even in the absence of soil pore evaporation, stomata do not shut down entirely at very high VPD so that ET does not go to zero, because stomata are still slightly open to perform some photosynthesis \citep{Ball_1987, Leuning_1990, MEDLYN_2011}. In addition, upward xylem transport is necessary to maintain phloem transport, as well as nutrient transport and thus carbon allocation \citep{De_2013, Nikinmaa_2013, Ryan_2014}.

\begin{table}  
\caption{Values of $VPD_{crit}$, where $\frac{\partial \; ET}{\partial \; VPD} = 0$, evaluated at PFT average values for $R_{air}$, $\gamma$, and $c_a$. PFT-specific constants ($g_1$, uWUE) are provided in Table \ref{pft}. For values of $VPD$ less than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be negative, and for values of $VPD$ greater than $VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}$ will be positive.}
\centering
\begin{tabular}{l c c c c c}
  \hline
  PFT & $R_{air}$ & $c_a$ (ppm) & $\gamma$  & \textbf{$VPD_{crit}$ (Pa)} \\
  \hline
  \input{vpd_crit.tex}
\hline
\end{tabular}
\label{vpd_crit}
\end{table}

Differences in $VPD_{crit}$ are exclusively determined by uWUE and the slope parameter ($g_1$) related to the plant functional type. A larger uWUE means a smaller $VPD_{crit}$, and an ET response to increases VPD that is more likely to be positive. At first glance this result is somewhat counter-intuitive; we expect that plants with a higher water use efficiency would be more water conservative. However, in reality uWUE determines how $WUE$ changes with VPD:

\[WUE = \frac{GPP}{ET} = \frac{uWUE}{\sqrt{VPD}}.\]
\[\frac{\partial \; WUE}{\partial \; VPD} = -\frac{uWUE}{2 \; VPD^{3/2}}\]

So, plants with a higher uWUE will have a greater \textit{decrease} in ecosystem-scale $WUE$ in response to increases in VPD. This decrease in $WUE$ causes more water loss per unit carbon gain, and explains the relationship between high uWUE and high likelihood of increases of ET in response to increasing atmospheric drying (increases in VPD).

A tendency towards increasing ET response with increasing VPD can also be caused by a high slope parameter ($g_1$), characteristic of plants that at the leaf scale are more willing to trade water for access to atmospheric CO$_2$. Plants that are less conservative will be thus be more likely to increase ET with increasing VPD. Both the aforementioned effects (large uWUE, $g_1$) can amplify each other, and generally conspire to shift the sign term towards a positive value for a given PFT.

This effect of uWUE and $g1$ on the sign term is most apparent by comparing two extreme PFTs: water intensive crops (CRO) and water conservative closed shrub (CSH). CRO has higher slope parameter and a slightyly higher uWUE ($g_1 = 140.7$ Pa$^{1/2}$; $2.85$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]) compared to CSH ($g1 = 75.1 \, Pa^{1/2}$, $uWUE=2.82$ $\mu$-mol [C] Pa$^{0.5}$ J$^{-1}$ [ET]). These differences in PFT parameters cause opposite ET responses to changes in VPD between CRO and CSH. ET theoretically always decreases with increasing VPD for the more water conservative CSH, while ET frequently  increases with increasing VPD for the more water intensive CRO (Figure \ref{idealized_sign}). CRO evolved or were bred to prioritize GPP and yield and are thus not water conservative. They are very willing to trade water for photosynthesis and productivity, despite changes in VPD, while CSH are very unwilling to trade water for more photosynthesis.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{./idealized_sign.pdf}
\caption{The functional form of the sign term, with $\sigma$ held fixed at 1, and all terms except VPD set to PFT averages. For comparison, the observed range of VPD for each PFT is plotted below the x-axis. Stars denote 25th, 50th, and 75th percentiles, and the range of the line spans the 5th-95th percentiles of observed VPD. Vertical black lines denote the location of $VPD_{crit}$ for each PFT, with the exception of CSH, for which $VPD_{crit}$ is off-scale.}
\label{idealized_sign}
\end{figure}

As expected, the slope parameter ($g_1$) is a primary determinant of the VPD dependence for the sign term shown in Figure \ref{idealized_sign}. Plants that are more conservative (small $g_1$) will tend to reduce ET with increasing VPD, and will be very effective at reducing ET, especially at low VPD. However, at very high VPD, gradients in vapor pressure at the leaf scale will become very strong as stomata reach their limits of closure in response to VPD (parameterized with $g_1$). As a result, ET response will begin to asymptote towards a constant ecosystem-scale values as leaf-scale response to VPD asymptotes towards zero.  Therefore, plants with a low $g_1$ will have the largest VPD dependence of ET response because the difference in ET response at low VPD (leaf stomatal response dominates) and high VPD (VPD gradient dominates) is largest. This is apparent in the strong VPD dependence of CSH, which has the lowest slope parameter ($g_1=75.1$ Pa$^{1/2}$) (Figure \ref{idealized_sign}).

To summarize our theoretical insights (Figure \ref{idealized_sign} and Table \ref{vpd_crit}), CROs are the least water conservative and have the strongest overall tendency to increase ET with increasing VPD, while CSH are the most water conservative and have the strongest tendency to decrease ET with increasing VPD, as well as the strongest VPD dependence of response. Figure \ref{idealized_sign} clearly shows, according to our theory, that for all PFTs except for crops there is frequent occurrence of a negative (plant dominating) ET response to increases in VPD. Therefore, plants are able in most atmospheric conditions to reduce ET in response to increased VPD and thus to reduce water loss. To better illustrate this, the ranges of observed environmental VPDs at the FLUXNET sites are plotted parallel to the x-axis. For CSH, VPD is always less than VPD$_{crit}$ (off scale) so that the plant response dominates in typical environmental conditions, emphasizing the water conservative strategy of those plants. For CRO on the other hand, VPD is higher than VPD$_{crit}$ for more than 50\% of observations, emphasizing that those plants operate with an aggressive water usage strategy, are water intensive and were actually engineered for photosynthesis rather than water saving. For DBF, EBF, MF, GRA, and SAV  more than half of the observed VPD are less than VPD$_{crit}$, i.e. in conditions where plant response dominates. SAV has a more water conservative response than the forest, grass, and crop plan types, but still respondes by increasing ET with increasing VPD for about a quarter of observations, due to the high aridity (VPD) of the SAV ecoclimate. It is also important to note that for all PFTs, even when atmospheric demand dominates, ET response to VPD is still far more negative than it would be for potential evaporation $\partial PET/\partial VPD$, i.e. atmospheric demand only, emphasizing that there is still a strong regulation of evaporative flux by stomata and though the plant xylem. The sign term in the PET case would just be a constant ($\frac{c_p}{R_{air}} \approx 3.5$), which is far larger than any part of the curves for any PFT. This highlights the deficiencies of PET's ability to capture land response to changes in atmospheric dryness (VPD). Plants are always regulating water exchange from the land surface, even when they reach the limits of they ability to do so.

\subsection{Functional Form of the Scaling Term}
\label{scale_term}
While the above discussion of the sign of $\frac{\partial \; ET}{\partial \; VPD}$ is important to answer our question of when ET response increases or decreases with VPD, understating the overall magnitude of the ET response is important to soil-plant-atmosphere water budgeting. So we now more closely examine the terms that affect how the sign term is scaled:

\begin{equation}
  \frac{g_a \; P}{T(\Delta + \gamma)}:
\end{equation}

$\frac{P}{T}$ is an air-density term, which varies little compared to aerodynamic conductance and Clausius-Clapeyron ($\Delta$). The psychometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clausius-Clapeyron relationship $\Delta$. This is as expected, given that the aerodynamic conductance represents the efficiency of exchange between the surface and the atmosphere. As aerodynamic conductance  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).

$\Delta$'s presence in the scaling term also matches physical intuition. $\Delta$ (and also the approximately constant $\gamma$) control the efficiencies with which surface energy is converted to latent and sensible heat \citep{Monteith_1965}. The functional from of $\Delta$ will be the same across PFTs, but the temperature range may vary slightly. In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness for aerodynamic conductance. So most of the differences in scaling between PFT should be in the aerodynamic conductance term.

The control of the scaling term variability between PFTs by aerodynamic conductance is confirmed by data (Figure \ref{scale_vary}). Differences between PFT are almost entirely due to differences in aerodynamic conductance, rather than differences in observed temperature ranges. The scaling term for the tree PFTs (DBF, EBF, ENF, MF) is generally about double the scaling terms for other PFTs which have lower surface roughness and generally smaller aerodynamic conductance (GRA, CSH, CRO). The savannah (WSA, SAV) PFT's scaling is somewhere betwen GRA, CSH, and CRO, and DBF, EBF, ENF, and MF, due to higher variability and surface roughness.

\begin{figure}
%\centering
\centerline{\includegraphics[width=1.4\textwidth]{./idealized_scale.pdf}}
\caption{Primary sources of variability for the scaling term, as a function of PFT. The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT.}
\label{scale_vary}
\end{figure}

Within each PFT, the scaling term variability is controlled both by environmental temperature and aerodynamic conductance variability (Figure \ref{scale_vary}). While the observed variability of the aerodynamic conductance contributes more to the scaling term variability than temperature, the temperature contribution is non-negligible. Specifically, the scaling term is generally larger at low temperatures when latent heat is relatively inefficient at moving energy away from the surface. This effect amplifies the role of aerodynamic conductance variability at low temperatures.

To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conductance, due to differences in surface roughness between each PFT, and possibly to a lesser extent wind conditions. In contrast, variability within PFT is also controlled by temperature, through Clausius-Clapeyron. But, aerodynamic conductance variability generally impacts the scaling term more than temperature, even within PFT.

\subsection{Bulk statistics of ET response to VPD}
\label{stats_sec}

In this section we consider direct observations of ET response with eddy-covariance data, while including uncertainty with the $\sigma$ term (Section \ref{methods}). These observational results of ET response largely confirm our theoretical analysis, presented above (Table \ref{stats}). For all PFTs, mean ET response to increasing VPD is negative. However, ET response evaluated at the average of all variables (e.g. $\sigma$, $T$, $c_a$, $VPD$) is positive for CRO, and negative for all other PFTs. This difference in mean ET response as compared to the ET response at mean environmental conditions is due to the non-linear nature of the response, in which negative responses are generally larger magnitude than positive responses (Figure \ref{idealized_sign}). Therefore, both the mean ET response as well as the ET response at mean environmental conditions matches our expectations from the theory (Section \ref{sign_term}), with the exception that CRO observations are shifted more towards a negative ET response than we expect.

\begin{table}
\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT.}
\centering
\begin{tabular}{l c c c}
  \hline
PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
  \hline
  \input{stats.tex}
  \hline
\end{tabular}
  \label{stats}
\end{table}

Regarding the frequency of negative and positive ET response, all PFTs exhibit a  decreasing ET response with increasing VPD (physiologically controlled, water conservative response) for the majority of observations. When listed by increasing frequency of negative response, the PFTs are generally ordered as we would expect them to be, from most water intensive to most water conservative: CRO (53\%), DBF (61\%), EBF (67\%), GRA (69\%), ENF (74\%), and CSH (97\%).

In general the bulk statistics match our theoretical expectations well, with the caveat that inclusion of uncertainty shifts crops towards a slightly more negative response than the theory and Figure \ref{idealized_sign} would suggest. The bulk statistics motivate a more thorough examination of the structure of uncertainty and more sophisticated validation of our theory's performance against observations.

\subsection{Validation of theory at eddy-covariance sites}
\label{testing}
We now compare more sophisticated distributions of the observed response to our simplified theory (Section \ref{sign_term}). The observed distribution of the sign term, as compared to what the theory would predict, is provided in Figure \ref{test_sign}. Our goal was to capture the leading order behavior of the ET dependence on VPD. Given the assumptions we have made, and the uncertainties of flux tower observations themselves, we expect a relatively large amount of noise when reproducing the derivatives of ET. However, the data largely reproduces our theoretical analysis.

\begin{figure}
\centering
\centerline{\includegraphics[width=1.4\textwidth]{./test_sign.pdf}}
\caption{Comparison of the sign term with model uncertainty included (box plots) to the sign term as calculated with simplifying assumptions (blue line, as in Figure \ref{idealized_sign}). Each box plot corresponds to 5\% of the data, and the 5-95\% range of VPD is plotted.}
\label{test_sign}
\end{figure}

For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the observations match the theory with the addition of noise. The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each bin. Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations. Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which comes from the observations themselves. The uncertainty is non negligible; there are many observations in each bin for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations well.

While CSH has a much different functional form of the sign term than DBF, CSH observations also match our theory to leading order. Again, the VPD dependence mostly determined from the slope parameter ($g_1$) closely matches the medians in the observation bins and the theoretical threshold is well reproduced (albeit off-scale). The VPD-independent, strongly negative response, which is determined by uWUE, is also captured. For CSH, there is extremely rare occurrence of observed positive ET response with VPD ($\approx 3$\%), even with uncertainty, so the sign of the observations almost always matches the sign of the theory, which states that ET response should always be negative.

As Table \ref{stats} suggested, the theory does not match as well for CRO as for DBF and CSH. At low VPD the VPD dependence (due to $g_1$) is similar to the leading order behavior of the observations, but the theory's curve is shifted towards more positive values relative to the observations, likely due to some bias in uWUE. Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependence, is behavior that is impossible for the theory to capture. This general positive bias of the theory explains the shift towards negative ET response observed in the bulk statistics in Table \ref{stats}. The divergence between the theory and the observations for CRO could have been expected. For practical reasons, our theory does not include important sources of plant type heterogeneity within and across sites for CRO. In agricultural landscape, the footprint of an eddy-covariance towers tend to not only observe a single species but many different ones, with potentially varying water use strategies. We also do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis and water strategies, or differences in irrigation practices between the sites and years. These deficiencies should largely explain the inability of our theory to exactly match the observations in croplands. For example, a superposition of sites with C3 plants (at low environmental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET response at high VPD when all sites are binned together, as in Figure \ref{test_sign}.

The theory for GRA suffers from similar limitations as for CRO. GRA observations are characterized by a consistent trend back towards negative ET response at higher VPD, which the functional form of our theory is incapable of accounting for. As compared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similar logic as we did for CRO. GRA also tends to exhibit substantial within-PFT variability, such as C3 vs C4 photosynthesis, as well as strong seasonal variability in height and LAI. As with CRO, the trend back towards negative ET response at higher VPD could be explained by a superposition of sites with C3 vs C4 plants. Unfortunately, because detailed site-level data of C3 vs C4 plant type is not available, we could not bin CRO and GRA further by irrigation and photosynthesis strategy. However, we hypothesize that the theory would validate against observations much better when photosynthesis pathways are accounted for.

The comparison between the theory and observations for ENF and EBF is unique from the rest of the PFTs. The VPD dependence of the theory, determined by the slope parameter ($g_1$), is not as well captured by observations. However, the theory still captures the general behavior and magnitude of the response, with some increase in error at low and/or high VPD.  We note that we used published constants for $g_1$ \citep[from ][]{Lin_2015, Franks_2017}, which determines the VPD dependence, and that deviations should be expected across sites \citep{Lin_2015}. We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori results whenever possible in order to avoid over-fitting and to have a theoretical approach as independent to the observations as possible for further confirmation of the validity of the approach.

While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the response at non-extreme VPD for ENF and EBF. It is useful to also compare our theory, the observed uncertainty in the ET response, and PET, as PET is often used as a proxy for ET in hydrometeorological studies and drought indices. Again, as in our discussion of the theoretical results, the PET sensitivity to VPD will be constant at about 3.5. In Figure \ref{test_sign} the ET response to shifts in dryness is always far smaller than 3.5 in the observations, even with all uncertainty included. So the PET response to atmospheric dryness is outside the range of all observable responses. This casts doubt on PET's usefulness for capturing the leading order behavior of water budget and dryness response at the land surface, as is done in several drought indices. However our theory (Equation \ref{et}) could be used with published uWUE and $g1$ data, as an alternative to PET to estimating land response and dryness, and would more accurately account for the response to atmospheric drying.

\subsection{Observed ET response to VPD}

Most of the results presented so far focus on the sign term, so now we turn to observations of ET response with the scaling term included (Figure \ref{data_scatter}). Until now, in the interest of developing leading order intuition for ET response, and to be conservative in our acknowledgment of model and observational error, we've considered $\sigma$ variability to be a measure of uncertainty. An alternative viewpoint is that $\sigma \cdot uWUE$ represents spatial and temporal variability of uWUE, which may be expected within bounds \citep[see Table \ref{pft}, also ][]{Zhou_2015}. This is a less conservative view; some of the $\sigma \cdot uWUE$ variability will be due to model and observational error, so by viewing $\sigma \cdot uWUE$ as \textit{real} variability we run the risk of mistaking noise for signal. However, the advantage of this viewpoint is that, from \citet{Zhou_2014}, we have very high confidence that uWUE fit at the hourly timescale (as we do with $\sigma \cdot uWUE$) correctly captures the relationship between ET, GPP, and VPD, and that our form of PM introduces minimal error with its use of uWUE.

\begin{figure}
\centering
\centerline{\includegraphics[width=1.4\textwidth]{./data_scatter.png}}
\caption{Scatter plots of observed $\frac{\partial \; ET}{\partial \; VPD}$, including $\sigma$ variability, as a function of PFT, temperature and VPD. Please note differences in the colorbar scale.}
\label{data_scatter}
\end{figure}

Therefore, with the caveat that some of the signal presented in Figure \ref{data_scatter} may in fact be noise, we can interpret the observed distribution of ET response to VPD. In general, the observed response matches the intuitive theory. ET response to VPD shifts towards positive values as VPD increases (atmospheric demand dominating). CRO exhibit the highest occurrence of positive ET response, and the observations confirm that CRO are the most water intensive. CSH are the most water conservative, with a strong negative response. DBF, EBF, ENF, and GRA are also water conservative, but show some occurrence of positive ET response to VPD, particularly at higher observed VPD, as the theory predicts.

Figure \ref{data_scatter} also includes the impact of the scaling term. For a given VPD, the magnitude of ET response does not vary strongly with temperature, confirming that any impact of the scaling term on the magnitude of the response is primarily due to changes in aerodynamic conductance. Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer. While $\Delta$ controls the efficiency of energy conversion to latent heat, it appears this is a second order term, relative to $g_a$, for scaling ET response.

As in Figure \ref{test_sign}, Figure \ref{data_scatter} matches our expectation based on simplified theory. The sign term is most strongly scaled by $g_a$, and in general the sign term shifts towards positive values as VPD increases. The willingness with which a given plant type evolved to use water dictates the occurrence of positive versus negative ET response. Water conservative plants are highly effective at mitigating the effects of atmospheric demand, and can store water for later use by reducing ET in response to increasing VPD.

\subsection{Limitations of theory - very dry soil moisture conditions}
In formulating our theory with Penman Monteith, we implicitly did not account for very dry soil moisture conditions. For the majority of environmental conditions  observed at the eddy covariance sites used here, soil conditions were not extremely dry so that we could assume a constant uWUE and $g_1$. We posit that ecosystems will generally optimize to host plants living in conditions which they evolved for. However, in extreme conditions and drought scenarios soil water content (SWC) could become the limiting factor for ET response to VPD, which our theory does not account for. In addition, low soil moisture conditions themselves increase VPD through land-atmosphere feedback \citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006, Salvucci_2013, Gentine_2016, Berg_2016}.

We first note that in the data used low SWC conditions are very infrequent. Within our framework, any systematic bias due to the  failure to account for SWC's effects in very dry conditions should manifest itself in a functional relationship between $\sigma$, our uncertainty measure, and SWC, which is observed at the FLUXNET sites. Examining the relationship between $\sigma$ and SWC will test to what extent our theory breaks down in very dry soil moisture conditions.

If we again view $\sigma \cdot uWUE$ as, in addition to a measure of uncertainty, a short time-scale observation of uWUE, we would expect $\sigma \cdot uWUE$ to decrease at low SWC. If $\sigma \cdot uWUE$ is a very strong function of SWC, then our theory should be conditioned more strongly on well watered soil conditions. If $\sigma \cdot uWUE$ is weakly a function of SWC, then our theory is more universal and independent of soil moisture conditions.

Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure \ref{swc_boxplot}). The portion of observations for which our theory is biased by SWC-limitations varies by PFT, due to the nonlinear threshold at which soil moisture availability limits plant function. For CRO and DBF, soil water content only matters for about the lowest 5\% of observations (each box is 5\% of observations in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very weak dependence on soil moisture, which is reflective of the high likelihood that CRO sites are optimally irrigated and not water stressed. For ENF and EBF, systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the lowest 30\% of SWC conditions. In crontast to ENF, CRO, and DBF, for GRA the relationship between $\sigma \cdot uWUE$ and SWC is more linear and affects a greater portion of observations; about 60\% of observations. Clearly, for GRA soil water frequently impacts plant function and alters ET response, and our theory is limited. It is therefore not suprising that our theory tested poorly againt the data for GRA, relative to ENF, EBF, DBF, and CSH (Section \ref{testing}). For all other PFTs occurence of soil mosture impacts was rare enough to not manifest itself in bulk statistics and plots. A possible exception is CSH, for which there are only two sites, which precludes comment on the exact PFT-wide relationship.

\begin{figure}
\centering
\centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box plot represents 5\% of all observations.}
\label{swc_boxplot}
\end{figure}

The observed dependence of $\sigma \cdot uWUE$ on SWC for GRA would explain the deficiencies of our theory compared to the observations in Figure \ref{idealized_sign}, specifically the trend back towards negative soil moisture at high VPD. Aforementioned feedbacks between the land surface and the atmosphere, which are not accounted for due to our focus on the one-way response of the land surface to atmospheric conditions, would cause high VPD to be correlated with low SWC \citep[][]{Gentine_2016, Berg_2016}. So, at high VPD observations of low SWC are more likely, and this low SWC causes a lower uWUE (Figure \ref{swc_boxplot}). The lower uWUE at low SWC/high VPD then leads to the observed downturn towards decreasing ET with increasing VPD, and the deviation between our theory (based on a constant uWUE assumption) and observations.

To summarize, our theory is limited by its inability to account for soil water impacts on land surface response, and feedbacks between SWC and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively rare (<30\% of observations) and does not manifest itself in the majority of observations and bulk statistics. However, for GRA SWC has a strong effect decreasing water use efficiency for the majority of the observations. Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA.

\section{Discussion}

We derived a new form of Penman Monteith using the concept of semi-empirical optimal stomatal regulation \citep{Lin_2015, MEDLYN_2011} and near constant uWUE \citep{Zhou_2015} to remove the implicit dependence of stomatal conductance on GPP and ET. With our new form of Penman Monteith we developed a theory for when an ecosystem will tend to reduce or increase ET with increasing VPD, which we evaluated against a range of eddy-covariance data spanning different climates and plant functional types. The goal was to capture the leading order behavior of the system to gain some intrinsic knowledge for its behavior. This intuition can be used to disentangle land atmosphere feedbacks in more complicated scenarios, and will aid interpretation of observations and sophisticated models.

Our theory suggests that for a majority of environmental conditions, plants will tend to conserve water and reduce ET with increasing VPD. Stomatal regulation and plant physiological response strongly regulate ET, and this regulation varies by PFT. CROs are the least water conservative, while DBF, EBF, ENF, GRA and CSH are progressively more water conservative (more likely to reduce ET in response to increasing VPD). Observations of ET response to VPD exhibit the same general behavior as the theory, with ET response becoming more positive (atmospheric demand dominating) as environmental VPD increases within a PFT, and more negative for PFTs that are adapted to arid conditions and prioritize water conservation over primary production.

The consequences of our theory our twofold. First, the ET response to VPD is always far more negative than it would be if only atmospheric demand is considered and plant physiological response is ignored (as with PET). Therefore,  any drought indices or analysis using PET will not capture how  vegetated surfaces respond and react to increases in atmospheric demand. However, the new uWUE-version of Penman Monteith we derived (Equation \ref{et}) could be used as a complement to PET in drought indices. One could calculate and compare indices  when plant response is ignored (PET) or included (Equation \ref{et}).

Lastly, our paper builds important intuition for how plants will respond to global change-induced shifts in VPD, which should increase in the future. Plant physiological responses to direct CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention than physiological response to indirect effects like increased VPD. Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (atmospheric drying). Additionally, we derive an explicit conceptual framework for ET response as a function of two plant type parameters, $g_1$ and uWUE (e.g. Equation \ref{et}). This framework can be used to understand ET response to a range of environmental change, and forms a foundation for resolving varying response to more nuanced consideration of plant type and climate. Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD). We argue that such simplified conceptual frameworks are critical tools for disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.

\acknowledgments
This work used eddy covariance data acquired and shared by the FLUXNET community, including these networks: AmeriFlux, AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont, ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC, OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance data processing and harmonization was carried out by the European Fluxes Database Cluster, AmeriFlux Management Project, and Fluxdata project of FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux offices. This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No. DGE 16-44869. All code and data used in this analysis, including those used to generate the figures and tables, is publicly available at https://github.com/massma/climate\_et.

\bibliography{references.bib}

\end{document}