### 2018-05-25 doc/paper/vpd_et_paper_arxiv.tex

\documentclass[12pt]{article}
\input{def.tex}

\begin{document}

\title{When does vapor pressure deficit drive or reduce
  evapotranspiration?}

\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
\author[1]{Pierre Gentine}
\author[1,2]{Changjie Lin}

\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
  Engineering, Tsinghua University, Beijing, CN 100084}

\maketitle

\begin{abstract}
  Increasing vapor pressure deficit increases atmospheric demand for water, and vapor pressure deficit is expected to rise with increasing greenhouse gases. While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water. Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET. We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.

 The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD. Plants that are evolved or bred to prioritize primary production over water conservation (e.g. crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing). However for forest, grass, savannah, and shrub plant types, ET more frequently decreases than increases with rising VPD. This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world. 
\end{abstract}

\onehalfspacing

\input{body.tex}