### 2018-10-30 doc/paper/ms.tex
forgot to add archive base file (ms.tex), updated readme

% this is used for arkiv submissions
\documentclass[12pt]{article}
\input{def.tex}

\begin{document}

\title{When does vapor pressure deficit drive or reduce
  evapotranspiration?}

\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
\author[1]{Pierre Gentine}
\author[1,2]{Changjie Lin}

\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
  Engineering, Tsinghua University, Beijing, CN 100084}

\maketitle
\begin{abstract}
  \input{abstract}
\end{abstract}

\onehalfspacing

\section{Introduction}

\input{intro.tex}

\input{body.tex}

\section{Conclusions}

\input{conclusions.tex}

\acknowledgments This work used eddy covariance data acquired and shared
by the FLUXNET community, including these networks: AmeriFlux, AfriFlux,
AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
are provided by ECMWF and processed by LSCE. The FLUXNET eddy covariance
data processing and harmonization was carried out by the European Fluxes
Database Cluster, AmeriFlux Management Project, and Fluxdata project of
FLUXNET, with the support of CDIAC and ICOS Ecosystem Thematic Center,
and the OzFlux, ChinaFlux and AsiaFlux offices. We would like to thank
Dr. Trevor Keenan for providing tools for citation of the FLUXNET2015
dataset, available at
\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material is
based upon work supported by the National Science Foundation Graduate
Research Fellowship under Grant No. DGE 16-44869. All code and data used
in this analysis, including those used to generate the figures and
tables, are publicly available at \url{https://github.com/massma/climate\_et}

\begin{appendices}
  \section{FLUXNET2015 sites}\label{flux_sites}
  % below taken adapted from Trevor Keenan's FLUXNET citations
  % at github.com/trevorkeenan/FLUXNET_citations
  Table A1: Metadata and citations for flux sites used in this analysis. All data are gathered from \url{www.fluxdata.org}, and citations are aggregated using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.
  \begin{longtable}{l l l l l l l}
    \hline
    \textbf{Site} &
    \textbf{PFT} &
    \textbf{Lat} &
    \textbf{Lon} &
    \textbf{Clim\textsuperscript{1}} &
    \textbf{Period} &
    \textbf{References} \\
    [0.5ex]
    \hline
    \input{flux_sites.tex}
    [1ex]
    \hline
  \end{longtable}
  \textsuperscript{1}  K{\"o}ppen Climate classification.
\end{appendices}

\bibliography{references.bib}

\end{document}