### 2018-12-17 doc/paper/hess-revision/review-1.tex
rough draft of RC1 response

\RequirePackage{natbib}
\bibliographystyle{agufull08}

Thank you very much for an excellent and very thoughtful
review. Addressing your comments will greatly improve the
manuscript. We agree with your comments, and think the manuscript
could use substantial reframing and rewording to clarify how we are
answering our research question, and how the answer may vary with
climate and environmental factors.

Just a quick general comment before addressing your specific
comments. By including a lot of content on PFT-focused analysis we
failed to communicate the major goal and scope of our project: we were
trying to characterize the response of ET to VPD, with all other
environment variables held fixed. To accomplish this, we needed to
formulate an explicit function of ET in terms of environemntal
variabiles and parameters, where any parameters can be approximated as
constant with regards to some [arbitrary] VPD perturbation
scenario. We will elaborate more below on why we used PFT-focused
scenarios for much of our analysis. However, the goal of our
manuscript was much more simple and fundamental: at a given place or
time, if you introduce a perturbation to VPD, what is the immediate ET
response (postive or negative)? Answering this question does not
require the much stronger assertion that any parameters must be
invariant within a given PFT.

\section{Compounding uncertainties}

We agree that the uncertainties are large both within a PFT and across
a PFT. Looking back, we believe that some of our language
communicating \cite{Lin_2018}, \cite{Medlyn_2017}, and
\cite{Zhou_2014}'s results was misleading, and this was exacerbated by
our focus on PFT-analysis. We will remove that language from the
manuscript, and add language to better communicate our results as
reflections of the considerable within-PFT uncertainty.

I think comparing the rankings of given PFT values for g$_1$ and uWUE
can be misleading, given the magnitude of the uncertainties
involved. Most of our calculated values for uWUE are within one
standard deviation of \cite{Zhou_2015}'s results, but these deviations
can result in some changes in the ordering of PFTs from high to low
uWUE. The bigger problem in our eyes is that we made a mistake with
some language suggesting within intra-PFT variability is less than
inter-PFT variability, but clearly this is not the case. We will
remove this language. These misinterpretation should not have made the
manuscript; we actually included \cite{Zhou_2015}'s results in Table 2
to be transparent about the within-PFT uncertainties. g$_1$ values
also exhibit considerable uncertainty, and again, while the relative
rankings of these values may change, they are all within observed
ranges in \cite{Medlyn_2017}. We did not include these in Table 2
because \cite{Medlyn_2017} does not provide the numerical values;
however, we think it would be useful to provide estimates of these
ranges form \cite{Medlyn_2017}'s figures in our Table 2.

Regarding how to interpret g$_1$ and uWUE: we think the most best way
is with established physical relations from \cite{MEDLYN_2011}
(Equation 3 in the manuscript) and \cite{Zhou_2014} (Pg. 7, line 10 in
the manuscript), with the knowledge that there is uncertainty
involved. The quantities in these relationships all have some
intrinsic physical meaning, but can vary substantially within a
PFT. Again, we need to alter our language to better reflect this
intra-PFT variability.

However, \cite{Zhou_2014} did establish a constant uWUE approximation
as a good approximation for capturing the relationship between GPP, ET
and VPD at a given place and time, so it is still a very useful and
robust approximation for answering our research question (see
discussion earlier in this comment). Additionally, a constant g$_1$
approximation is used in many earth system models, so while it
introduces considerable uncertainty, it also makes our framework
useful for interpreting modelled vegetation response in ESMs and GCMs.

We also agree with the comment that the best way to think of this is
probabalistically. This is the most robust approach to dealing with
approximations - we introduce randomness to our variables and
paramters to account for all of the physics that are not explicitly
accounted for, as well as observational uncertainty. However, as far
as I know we still have not developed a general, robust, and efficient
arithmetic for random variables. We could try and adapt a Bayesian
model to this problem, but given the large amount of arithmetic
invloved, fully incorporating a Bayesian represntation to every
variable in the analysis would be a very hard problem, and a
significant research project in its own right. I think a good
compromise is to add language directing the reader to interpret our
results more probabalistically, which we have already presented
probablistically in the figures. For example, in Figure 5 the range
of values in each plot represents a range of possible responses in the
sign term. Some portion of this could be due to oveservational
uncertainty, and others could be due to physical intra-PFT variations
inducing changes in uWUE and/or g$_1$. It's also worth noting that our
intra-PFT variability is greater than the inter-PFT
variability. Again, we need to add language highlighting this, and its
consistency with previous results. We included this uncertainty
explicitly to be transparent.

\subsection{Attribution to physiological responses}

We do not want our analysis to be interpretted as an assertion that
g$_1$ and uWUE are attributes of PFT only. We expect them to vary both
within a PFT and across a PFT. Our primaty goal in using them was to
develop an explict expression of ET as a function of environmental
variables, and use this to assess the response to a change in VPD. We
focused our analysis using PFTs because this is how it was framed in
previous studies in our field, and specificically the studies used in
our derivation \citep{Zhou_2014, Zhou_2015, Medlyn_2017}. We were
thinking the PFT-focused analysis could be useful, especially given
that climate models generally hold plant paramters fixed with respect
to PFT, so long as we were transparent about the large uncertainties
and problems with this approach (see Figure 5).

We agree with you, and think stating that any quantity is fixed within
a PFT is hard to believe. Phenotypic variation and adaptation within a
given specicies can be considerable (i.e. effecting $\lambda$, g$_1$,
and uWUE), so it would be hard to say that anything would be constant
within a PFT made up of many different diverse species. Both
phenotypic variation as well as the species distirbution and dominant
PFT at a given location will all be strongly optimized in response to
climate. In this sense, the distribution and evolution of local
climate is a strong control on the local ecosystem structure and
physiology of that ecosystem.

What we need to communicate better is that a given ecosystem's state
at a place or time controlls its response to a VPD perturbation. We
are making an approximation that we can parameterize the effect of the
ecoystems' state on VPD response with g$_1$ and uWUE. These quantities
can also vary due to soil moisture condition; the approximation we
need to make to answer for our research question is that they are
fixed with respect to a VPD perturbation. In this sense, we do not
view the two statements: ``plants that are evolved to bred to
prioritize primary production over water conservation (e.g., crops)
exhibit a higher likelihood of atmospheric demand-driven response''
and ``ecosystem types are responding in this way because they have, on
the whole, been subjected to less soil water limitation (due to the
non-negligible effects of irrigation?)'' as mutually exclusive. In
fact, we view them as consistent with a view that crops and thier
physiology (parameterized by g$_1$ and uWUE) exist in a given
environment because they have not been subjected to soil water
limitation, and they have a given response to VPD because of their
physiology (which is a direct effect of the environment). In this way,
climate and land surface state are causes of the VPD response through
their effect on plant physiology (parameterized by g$_1$ and uWUE).

When writing this manuscript, there was definitely some internal
tension and debate about how to frame the analysis and results. We
could either focus on PFT-oriented results as previous literature has
done (e.g. holding plant physiology fixed within a given PFT), or
allow plant physiology terms to vary through time and space and look
at the distribution of ET response to VPD (see Section 3.5). After
re-examining the manuscript both after some time away from the
problem and in light of your comments I think a strong argument could
be made that we made the wrong choice with respect to this focus. It
may have made more sense to focus our analysis on the ET response more
generally across space and time as ecosystem-scale plant physiology
varies in response to climate and soil moisture.

\section{Next Steps}

In order to fix the manuscript and our communication of the answer to
the question ``When does VPD drive or reduce ET?'', we see a few
potential paths that we will consider between now and the final
response after the discussion period. If you (or anyone else) has any
opinions or comments in the meantime, we would appreciate the insight
and feedback.

\begin{enumerate}
\item We include the discussion presented in this review and response,
  and include language explicitly stating that the purpose of the PFT
  analysis is to provide connections to other PFT-constant analyses
  and models (e.g. ESMs and GCMs). We will add extensive language on
  the uncertainty and weaknesses of this approach, and reframe our
  conclusions to reflect his uncertainty.
\item We instead alter our analysis to look at how VPD response
  varies with climate-induced variations and general plan
  physiological variability instead of just PFT variability that
  poorly captures all of the observed variability in
  response. Sections 3.1 - 3.2 would be replaced by analysis direcly
  relating ET response to general climate and plant physiological
  terms, rather than PFT-mean analysis. For example, the ``scaling
  term'' analysis would be presented in terms of generic changes in
  plant height and temperature, and the ``sign term'' analysis would
  be framed by generic changes to g$_1$ and uWUE, as informed by
  previous literature.
\item This is the most extreme option in terms of modifying the
  manuscript. We significantly alter the manuscript and instead focus
  on just the general shape of the ET-VPD curve (with environmental
  variables held fixed). This was not discussed in the review, but one
  of our most noteworthy results is on the general shape of the ET-VPD
  curve being concave up, given an assumption of a square-root VPD
  dependence of plant response. This result is independent of any
  assumptions of uWUE and g$_1$, and to our knowledge it is first
  derived curve of ecosystem response to VPD (see Section 3.7). It
  also highlights the importance of discerning the exact exponent of
  VPD dependence, as this alters the fundamental nature and shape of
  the curve. Essentially, the manuscript would become just our
  motivation and derivation, followed by Section 3.7 exploring the
  consequences of the derivation for the shape of the VPD curve. This
  alteration of the manuscript would likely result in more of a
  technical note type paper, and we are hesitant to do this because we
  think there is still a lot of useful information obtained by tying
  our results to real-world scenarios (as in Options 1 and 2 above).
\end{enumerate}

Here we present some final minor comments and concerns on the
technical details of our proposed changes to the manuscript, which are
relevant to representing uncertainty and spatiotemporal variabiliy in
uWUE \textit{and} g$_1$. In the original manuscript we used a single
$\sigma$ term to represent this variability. We did this because there
was changes in uWUE and g$_1$ induce a very similiar change in the ET
solution, which made solving for independent $\sigma_{uWUE}$ and
$\sigma_{g1}$ terms intractable, and represnting variability in both
g$_1$ and uWUE difficult. We decided to hold g$_1$ fixed within a PFT
for a two reasons: 1) ESMs and GCMs generally hold g$_1$ fixed, and 2)
letting uWUE vary seemed more appropraite to represent specifically
soil water variations on stomatal conductance, as $uWUE$ modifies
stomatal conductance analgalously to how soil moisture factors modify
a maximum stomatal conductance in land surface models (e.g. it is a
multiplicative factor on the entire stomatal conductance
term). Because uWUE and g$_1$ induce similar changes in ET, at least
qualitatively we think that having a single $\sigma$ can at least
represent some of the variability in both $uWUE$ and g$_1$. However,
formally there is no explicit variability in the g$_1$ term. The
discussion in this review rightly points out that based on previous
results we do expect some variability in g$_1$ as well. So, an
unresolved question is how important is it to the analysis and its
interpretation to include an explicit g$_1$ variability term in
addition to the exisiting $\sigma$ term. I'm including these comments
to hopefully stimulate some discussion. I could imagine some timescale
based approaches where we might be able to account for both the g$_1$
and uWUE variability, for example by making assumptions over what time
scale each quantity is fixed, and fitting based on that (e.g. g$_1$ is
fixed for a given season, and uWUE is fixed for a given day). This
approach might allow a tractable solution, and also could help us
filter error and observational noise from ``true''  plant
physiological and climatic variability in uWUE and g$_1$. The cost of
all of this is a significant increase in the analysis and content of
the paper, as well as some increased opacity to the methods.

Thank you again for the thoughtfull review. We hope it stimulates
further discussion.

\bibliography{references.bib}