### 2018-12-19 doc/paper/hess-revision/review-2.tex
done with comment (ii)

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{xcolor}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{color}
\pagestyle{plain}

\graphicspath{../../../etc/plots/climate_et/test_plots}
\RequirePackage{natbib}
\bibliographystyle{agufull08}

% Bib aliases
\makeatletter
\def\@citex[#1]#2{\leavevmode
  \let\@citea\@empty
  \@cite{\@for\@citeb:=#2\do
    {\@citea\def\@citea{,\penalty\@m\ }%
\edef\magic##1{\let##1\expandafter\noexpand\csname bibalias@\@citeb\endcsname}%
\magic\tmp \ifx\tmp\relax\else \let\@citeb\tmp\fi
     \edef\@citeb{\expandafter\@firstofone\@citeb\@empty}%
     \if@filesw\immediate\write\@auxout{\string\citation{\@citeb}}\fi
     \@ifundefined{b@\@citeb}{\hbox{\reset@font\bfseries ?}%
       \G@refundefinedtrue
       \@latex@warning
         {Citation `\@citeb' on page \thepage \space undefined}}%
       {\@cite@ofmt{\csname b@\@citeb\endcsname}}}}{#1}}
\def\bibalias#1#2{\expandafter\def\csname bibalias@#1\endcsname{#2}}
\makeatother

\begin{document}

Thank you for a very thoughtful review; as in our response to RC1, we
feel very lucky to have received such quality reviews.

We read through your major and minor comments, and do not anticipate
issues reconciling your minor comments for the final response after
the discussion period. We will focus out response on your major
comments, with the goal of opening the door to further discussion.

Also please note our response to RC1, as some of the discussion there
is relelvant to this response; particularly discussion on PFT-scale
analysis \url{https://editor.copernicus.org/index.php/hess-2018-553-AC1.pdf?_mdl=msover_md&_jrl=13&_lcm=oc108lcm109w&_acm=get_comm_file&_ms=72556&c=153745&salt=17166479111051651323}.

\section{Major comment (i)}

One issue with analyses attempting to diagnose the ET response to VPD
with pure data (i.e. using binned boxplots, as you suggest), is that
ET varies strongly with both aerodynamic resistance and
radiation. Simply binning by VPD or by site includes too many
confounding factors to really diagnose what the attirbutable response
is to VPD. This issue is really what motivated our analysis and makes
it unique; by building an analytical framework based on confirmed
results in the literatre (uWUE, Medlyn model, Penman-Monteith), we are
able to formally deduce what the actual ET response is, with other
environmental factors held fixed. This ``other environmental factors
held fixed'' component is what is, in our mind, nearly impossible to
deduce from pure data, because the environment is always changing and
direct analogues are either impossible to find or result in too small
of a sample size to form a meaningful conclusion. However, while we
cannot assess our formulation of the derivative with respect to VPD,
we can assess the actual model of ET, which includes all
approximations we introudced (specifically uWUE). As a part of our
analysis test suite we generated a figure of this comparison and
reproduce it here (Figure \ref{rmse}). Apologies this is not a
publication quality figure, but we wanted to get this response out
sooner rather than later to stimulate discussion, especially as we are
already behind because of the AGU Fall Meeting. This comparison is
fair in the sense that by using uWUE we introduce an extra free
parameter, so in the original PM model we introduce an extra free
parameter modifying g$_s$ and fit it also for each given PFT. The
introduction of the uWUE approximation does degrade the quality of the
ET model relative to the original PM model. However, we expect this,
as using uWUE is a simplification over using GPP directly, and may
break down in extreme, limiting cases like VPD $\to$ 0. As
a purpose-built model for assessing the ET response, we think the uWUE
approach is justified. Essentially we are saying that to the degree we
trust arithmetic, uWUE, Penman-Monteith, and the Medlyn stomatal
conductance model, our evaluation of the derivative is robust.

\begin{figure}[h]
  \centering \includegraphics{./rmse_bias.png}
  \caption{Mean bias (absolute value) and root mean squared error
    (RMSE) for three different ET estimates, relative to FLUXNET-2015
    observations: original (original Penman-Monteith formulation,
    e.g. Equation 1 in the manuscript, with Equation 4 providing
    g$_s$; uWUE (Equation 8 in the manuscript); iWUE (as in Equation
    8, but if we had used iWUE to remove GPP dependence rather than
    uWUE).}
  \label{rmse}
\end{figure}

However, the discussion so far assumes that the variables in our model
(Equation 8) do not change directly with VPD and we are free to
evaluate the derivative. You bring up a very good example that we did
not consider: a change in VPD will induce a change in net radiation
through surface tmperature that PM does not consider. We will
definitely add language addressing this oversight. Ideally, we would
incorporate that effect directly in Equation 8. We briefly tried to do
this in time for this response, but difficulties deriving a diagnostic
equation for temperature sabatoged the effort, and we do not want to
delay the response further. We will continue to work on this for the
final response phase. In the meantime, we think we can make some
logic-based arguments in the interest of discussion as to what the
effect would be. In the case where the aerodynamic and physiological
ET response (which the manuscript does consider) increases with VPD,
we would expect a decrease in surface temperature. This would induce
an increase in net radiation, as OLR and ground heat flux
decrease. This increase in net radation would induce a further
increase in ET response. Granted, the net radiation perturbation would
damp the decreased temperature response to the
aerodynamic/physiological terms, but all of the signs of the change
would stay the same. The converse is true for a decrease in ET with
respect to VPD: surface temperature increases, OLR and ground heat
flux increase, net radiation decreases, ET decreases. In this way, we
believe the effects of VPD on radiation do not change the nature of
the sign of the response, but could amplify the magnitude. Comments on
this are incouraged, and we will work to include a more quantitative
and rigorous analysis in the final response. Sometimes it is easy to
trick onself with these types of qualitative arguments, particularly
with complicated systems, so hopefully we have not presented faulty
logic in our rush to make this response public.

\section{Major comment (ii)}

Thank you for having an open mind about the concave up ET-VPD curve
result and providing clear explanations of concerns about this result
based on literature. Given the apparent controversy, we will
definitely add more language connecting previous results, particularly
at the leaf scale, with the presented results. Here we will informally
elaborate in the interest of discussion.

One of the primary differences between ecosystem scale assessments of
ET (specifically using Penman Monteith) and experimental and modeling
results of leaf scale response to VPD is that by taking an energy
balance approach Penman Monteith accounts for changes in $e_{sat}(T)$
at the surface induced by ET. To demonstrate this, compare Figure
\ref{leaf} and Figure \ref{PM}. In Figure \ref{leaf} we apply a naive
leaf scale approach to calculate ecosystem ET without the energy
balance consideration introduced by PM. In this figure, the curve
shows a concave downward shape, in direct analogy to leaf scale
T. Once the effects of evaporation's thermodyanmics are included in
the energy balance, the curve's shape is no longer concave
downward. How these conceptual results translate to leaf scale
experiments depends on the experimental setup
\citep[e.g.,]{Rawson1988, Turner1984, Mott2013}; however generally
chamber based leaf scale experiments do not preserve the energy
balance relationships we expect for a free surface in a natural
environment (sometimes intentionally by design). Additionally, we
would like to note that a concave up result is not necesarily
inconsistent with the statement ``transpiration at the leaf scale is
mostly positive'' with VPD. Especially for plants in well watered
environments like crops, the manuscript's ecosystem scale shape
(Figure \ref{uWUE}) would not be out of place on the figures in
\cite{Rawson1977} and \cite{Mott2007} for the range of VPD considered
(Figure \ref{uWUE} compared to Fig. 2 in \cite{Rawson1977}, also
Figure 7 in \cite{Mott2007}). However, the change in relative
magnitudes are not similiar, and this effect can be traced partially
to the effect of decreasing surface $e_{sat}(T)$ in response to
increasing evapotranspiration. While not direcly considered here (or
in the manuscript) we aslo think it is important to consider how plant
physiology in the natural environment could vary from plants grown in
well-watered conditions in experiments. It is not inconceivable that
phenotypic variability and adaptation in response to water limitation
could result in plants with different responses to those grown in a
lab.

The previous paragraph discussed the change in frame of reference
induced by using PM to assess ecoystem scale ET rather than directly
applying leaf scale logic in controlled experiments. However, our
uWUE-based PM framework additionally accounts for changes in
photosythesis induced by VPD. Discussion in \cite{Damour2010} (``The
issue of co-regulation of g$_s$ and A$_{net}$'') provides some
interesting background both on the general weaknesses of current
stomatal theory, and specifically on the possible importance of resolving the
effects of water stress on $A_{net}$ in stomatal conudctance. Given
that the current state of the art in physically-derived theory has not
converged on the proper way to account for changes in $A_{net}$ and
$g_s$, we believe using the validated uWUE empiracle results
\citep[e.g. Figure 3]{Zhou_2014} is a good approach to develop an
estimate of ET response to VPD accounting for changes in
photosythesis. This method of using an empricle result in the absence
of a complete physical theory has been used often in science
(e.g. turbulence) to increase our knowlege about a given systems'
behavior, and we think it is well motivated here. However, that said
it is important to add language and consideration on how introducing
this empricism could alter interpretation of our results. In
particular, while uWUE was developed using leaf scale photosythesis
theory, it was validated with observations of a fully coupled
land-atmosphere system. It captures a relatively consistent observed
relationship between GPP, ET, and VPD as they vary. There is a chance
that some of this relatioship could be due to feedbacks between the
land and the atmosphere, even though the relationship was derived by
leaf scale theory without feedbacks (e.g., if the assumptions behind
the uWUE theory introduced ``real'' errors that were compensated for
by feedbacks or ecosystem-scale physics in the observations).  Were
this the case, then we introduced a conceptual impurity to the
statement ``we are examining the one way response of ET to VPD.''
But, given the theoretical motivations for uWUE we think the portion
the relationship related to land-atmosphere feedbacks are likely
small, if they exist at all, and regardless this approach is still an
improvement over ignoring the documented effect of water stress on net
photosythesis.

Given the continued gaps in our ability to represent stomatal behavior
\citep(Damour2010}, particulary for ecosystems in natural
environments, we think it is hard to rule out our results. In the end,
it comes down to the degree to which we trust our three tools: the
Medlyn model, Penman-Monteith, and uWUE. To us these offered the most
compete analytical approach to the problem at present. However, it's
important to emphasize that we are not saying this is certainly
correct, and to do so would be irresponsible given the continued
development of stomatal theory and how it translates to the ecoystem
scale. What is important is to recognize that something as fundamental
as the shape of the ET-VPD curve is still very much an area of open
research, and sensitive to the representation of plant physiology (of
which there are many possibilities) in a given model or framework. We
thank the reviewer for having an open mind and giving us a chance to
elaborate; others in the community have rejected this result off-hand
(or even ridiculed it), which reinforces for us the importance of
demonstrating current uncertainty in VPD-ET relationships.

Just a quick couple of final notes:

We think that we miscommunicated with our (mis)use of ``plant physiology''
throughout the manuscript, and will alter our language in the final
review phase to be more precise. Many times when we refer to ``plant
physiology'' we mean any effect (direct or indirect) of an ecosystem
response, e.g. any behaviour that would not be observed over a wet or
bare soil surface.

\end{document}