### 2019-02-27 doc/paper/intro.tex

Vapor pressure deficit (VPD) is expected to rise over continents in
the future due to the combination of increased temperature and,
depending on region, decreased relative humidity
\citep{Byrne_2013}. Increases in VPD increase the atmospheric demand
for evapotranspirated water \citep{Penman_1948, Monteith_1965}, but
also stress plant stomata \citep{Leuning_1990, MEDLYN_2011}.

The opposing effects of increased atmospheric demand and higher
stomatal stress lead to two possible perspectives for how
evapotranspiration (ET) responds to shifts in VPD. The first, a
hydrometeorological perspective, is that higher VPD increases
atmospheric demand for water from the land surface, and this drives an
increase in evapotranspiration (ET). However, plants' stomata have
evolved to optimally regulate the exchange of water and carbon, and
tend to partially close in response to increased atmospheric dryness
\citep{Farquhar_1978, Ball_1987, Leuning_1990, MEDLYN_2011}. This
leads to a plant physiology perspective, in which an increase in VPD
may actually correspond to a decrease in ET because of stomatal
closure \citep[e.g.][]{Rigden_2017}.  In other words, the question
``When does VPD drive or reduce ET?'' can be related to whether plant
regulation or atmospheric demand dominates ET response.

The ET response to changes in VPD alters water partitioning between
the soil and atmosphere. If ecosystem plant response reduces ET with
atmospheric drying then soil moisture will be better conserved. This
represents a sensible evolutionary strategy to cope with aridity: save
water for periods when atmospheric demand for water is relatively low,
and atmospheric carbon can be accessed with a relatively smaller cost
in water loss. If instead stomata were fully passive \citep [similar
to soil pores, e.g. ][]{Or_2013}, increased atmospheric aridity would
strongly reduce soil moisture \citep{Berg_2017}. This could further
increase aridity as low soil moisture levels increases the Bowen
ratio, leading to increased temperature and atmospheric drying
\citep[][]{Bouchet_1963, Morton_1965, Brutsaert_1999, Ozdogan_2006,
  Salvucci_2013, Gentine_2016, Berg_2016}. Therefore, passive
regulation and a lack of soil moisture conservation does not seem to
be a sensible strategy for plants from an evolutionary standpoint.

We can use intuition about plant water conservation strategy to
hypothesize about ET response to changes in VPD. Plants and ecosystems
that evolved to conserve water, such as arid shrubs, should be more
likely to reduce ET with increasing VPD, and plants that have evolved
or have been engineered to care little about water, such as crops,
will be more likely to increase ET with increasing VPD. Atmospheric
conditions must matter as well. At the ecosystem scale, there are
limits to plant water conservation strategies. As atmospheric demand
for water (VPD) increases, ecosystems should begin to reach their
water conservation limits and might not be able to entirely limit ET
flux to the atmosphere. At this stage any further increase in VPD will
most likely drive a (limited) increase in ET, because the increase in
atmospheric demand for water overwhelms the limited plant response to
conserve water.

The objective of the present manuscript is to use reasonable
approximations established in prior research as a tool to develop a
framework for understanding plant response to atmospheric drying and
evaluating the VPD dependence of ET. This framework will aid
interpretation of observations, full complexity models, and facilitate
the disentanglement of complex land-atmosphere feedbacks. In the past,
similar approaches were used to understand interactions between
stomatal conductance, evapotranspiration and the environment
\citep[e.g.,][]{Jarvis_1986, Mcnaughton_1991}. However, at the time
researchers' understanding of the form of VPD's effect on plant
physiology was limited, so they could not explore the sensitivity of
ET to VPD, including VPD's effect on stomatal conductance and plant
function.

Recent results have drastically improved our understanding of VPD's
impact on physiology, especially at the leaf
level. \citet{MEDLYN_2011} developed a model for leaf-scale stomatal
conductance ($g_s$), including VPD response, by combining an optimal
photosynthesis theory \citep{Cowan_1977} with an empirical approach,
and extended use of this model to the ecosystem scale in
\citet{Medlyn_2017}. Additionally, \citet{Zhou_2014} demonstrated that
a quantity underlying water use efficiency $\left(uWUE = \frac{GPP\;
\sqrt{VPD}}{ET}\right)$ properly captures a constant relationship
between GPP, ET, and VPD over a diurnal cycle at the ecosystem
scale. uWUE is also relatively well conserved in the growing season
across space and time, within a plant functional type (PFT)
\citep{Zhou_2015}. While stomatal conductance parameterizations and
uWUE greatly simplify complex plant physiological processes, they
still capture ecosystem behavior for vegetated surfaces
\citep{Medlyn_2017, Zhou_2014}, and are novel tools to transparently
develop intuition for the behavior of complex, multiscale
ecohydrologic systems.

In this manuscript, we leverage uWUE and recent developments in
stomatal conductance parameterizations \citep{MEDLYN_2011} to derive
the theoretical one-way response of ET to VPD with other environmental
variables properly controlled for, i.e. we develop a framework for
evaluating the partial derivative of ET with respect to VPD. For the
first time, we explicitly include VPD's full effect on stomatal
conductance, including its impact on photosynthesis. We explore the
range of possible ET responses to VPD, given parameters previously
established in peer reviewed literature. Additionally, we show the
sensitivity of the ET-VPD relationship to model and framework choice,
highlighting the importance of: 1) future research on stomatal
conductance and ecosystem scale modeling, and 2) thoughtful selection
of photosynthesis and stomatal conductance models in more
sophisticated land surface and earth system models.