### 2019-03-01 doc/paper/hess-supplement.tex
added more supplementary material and it compiles

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{xcolor}
\usepackage{hyperref}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{color}
\pagestyle{plain}
\usepackage{longtable}
% \usepackage[
% 	backend=biber,
% 	url=false,
% 	isbn=false,
% 	doi=false,
%         eprint=false,
%         bibstyle=authoryear,
%         style=authoryear,
%         % style=numeric-comp,
% 	% uniquename=false,
% 	% uniquelist=false,
% 	% firstinits=true,
%  	natbib=true,
% 	sorting=none,
%         ]{biblatex}

% \addbibresource{./references.bib}

\RequirePackage{natbib}
\bibliographystyle{agufull08}
\maxdeadcycles=200

% Bib aliases
\makeatletter
\def\@citex[#1]#2{\leavevmode
  \let\@citea\@empty
  \@cite{\@for\@citeb:=#2\do
    {\@citea\def\@citea{,\penalty\@m\ }%
\edef\magic##1{\let##1\expandafter\noexpand\csname bibalias@\@citeb\endcsname}%
\magic\tmp \ifx\tmp\relax\else \let\@citeb\tmp\fi
     \edef\@citeb{\expandafter\@firstofone\@citeb\@empty}%
     \if@filesw\immediate\write\@auxout{\string\citation{\@citeb}}\fi
     \@ifundefined{b@\@citeb}{\hbox{\reset@font\bfseries ?}%
       \G@refundefinedtrue
       \@latex@warning
         {Citation `\@citeb' on page \thepage \space undefined}}%
       {\@cite@ofmt{\csname b@\@citeb\endcsname}}}}{#1}}
\def\bibalias#1#2{\expandafter\def\csname bibalias@#1\endcsname{#2}}
\makeatother

\title{Supplementary material: When does vapor pressure deficit drive or
  reduce evapotranspiration?}

\author{Adam Massmann, Pierre Gentine, and Changjie Lin}

\date{February 28th, 2019}

\begin{document}

\maketitle

The manuscript analyzes the partial derivative of ET with respect to
VPD. This assumes that all other quantites remain fixed, including the
plant parameters g$_1$ and uWUE (and by extension, $\lambda$). In
reality, these parameters may vary with environmental conditions, and
specifically soil moisture. However because soil moisture
\textit{only} enters the partial derivative directly through these
plant terms, if the plant parameters are weak functions of soil
moisture then the theory can be directly applied to a broader range of
conceptual VPD scenarios, including observed compound events between
high VPD and low soil moisture \citep{Zhou_2019}. To help the reader
assess the soil moisture dependence of uWUE (and partially by
extension $\lambda$ and g$_1$), we provide here figures showing the
distribution of uWUE with SWC for each of 66 FLUXNET sites from the
FLUXNET-2015 database. The functional relationship between uWUE and
SWC varies, with a mix of sites showing strong and weak
SWC-dependence. For all sites the ratio of signal to noise is very
low, an unfortunate consequence of taking a ratio of two highly
uncertain eddy-covariance derived fluxes. In general we find this
analysis inconclusive.

The paper does not rely on assumptions about uWUE's functional
relationship with soil moisture so we do not include the figures in
the body of the manuscript. But given that constant uWUE g$_1$
assumptions can make our theory more useful to the reader we provide
the figures for their interpretation, and motivation for future
research.

\input{supp-figs.tex}

\section{FLUXNET2015 sites}
  % below taken adapted from Trevor Keenan's FLUXNET citations
  % at github.com/trevorkeenan/FLUXNET_citations
  Table A1: Metadata and citations for flux sites used in this analysis. All data are gathered from \url{www.fluxdata.org}, and citations are aggregated using tools available at \url{https://github.com/trevorkeenan/FLUXNET_citations}.
  \begin{longtable}{l l l l l l l}
    \hline
    \textbf{Site} &
    \textbf{PFT} &
    \textbf{Lat} &
    \textbf{Lon} &
    \textbf{Clim\textsuperscript{1}} &
    \textbf{Period} &
    \textbf{References} \\
    [0.5ex]
    \hline
    \input{flux_sites.tex}
    [1ex]
    \hline
  \end{longtable}
\textsuperscript{1} K{\"o}ppen Climate classification.

\bibliography{references}

\end{document}