### 2019-06-20 doc/paper/abstract.tex
read through once, just final read through and submit

Increasing vapor pressure deficit (VPD) increases atmospheric demand
for water. While increased evapotranspiration (ET) in response to
increased atmospheric demand seems intuitive, plants are capable of
reducing ET in response to increased VPD by closing their stomata, in
an effort to conserve water. Here we examine which effect dominates
response to increasing VPD: atmospheric demand and increases in ET, or
plant response (stomata closure) and decreases in ET. We use
Penman-Monteith, combined with semi-empirical optimal stomatal
regulation theory and underlying water use efficiency, to develop a
theoretical framework for understanding how ET responds to increases
in VPD. The theoretical ET response to VPD over a range of reasonable
environmental conditions and plant characteristics varies from a
strong decrease in ET in response to increasing VPD (water
conservative) to a strong increase in ET in response to VPD (water
intensive), highlighting the diversity of plant water regulation
strategies.

The ET response varies due to: 1) climate, with tropical and temperate
climates more likely to exhibit a positive ET response to increasing
VPD than boreal and arctic climates; 2) photosynthesis strategy, with
C3 plants more likely to exhibit a positive ET response than C4
plants; and 3) due to plant type, with crops more likely to exhibit a
positive ET response, and shrubs and gymniosperm trees more likely to
exhibit a negative ET response. These results, derived from previous
literature connecting plant parameters to plant and climate
characteristics, highlight the utility of our simplified framework for
understanding complex land atmosphere systems in terms of idealized
scenarios in which evapotranspiration responds to VPD only. This
response is otherwise challenging to assess in an environment where
many processes co-evolve together.