### 2019-09-06 doc/paper/flux_sites.tex

AT-Neu & GRA & 47.1167 & 11.3175 & Unk & 2002-2012 & \cite{AT-Neu} \\
AU-ASM & ENF & -22.2830 & 133.2490 & Unk & 2010-2013 & \cite{AU-ASM} \\
AU-Cpr & SAV & -34.0021 & 140.5891 & Unk & 2010-2014 & \cite{AU-Cpr} \\
AU-DaP & GRA & -14.0633 & 131.3181 & Aw  & 2007-2013 & \cite{AU-DaP} \\
AU-DaS & SAV & -14.1593 & 131.3881 & Aw  & 2008-2014 & \cite{AU-DaS} \\
AU-Dry & SAV & -15.2588 & 132.3706 & Unk & 2008-2014 & \cite{AU-Dry} \\
AU-Gin & WSA & -31.3764 & 115.7138 & Unk & 2011-2014 & \cite{AU-Gin} \\
AU-How & WSA & -12.4943 & 131.1523 & Aw  & 2001-2014 & \cite{AU-How} \\
AU-Rig & GRA & -36.6499 & 145.5759 & Unk & 2011-2014 & \cite{AU-Gin} \\
AU-Stp & GRA & -17.1507 & 133.3502 & Unk & 2008-2014 & \cite{AU-DaP} \\
AU-Tum & EBF & -35.6566 & 148.1517 & Cfb & 2001-2014 & \cite{AU-Tum} \\
AU-Whr & EBF & -36.6732 & 145.0294 & Unk & 2011-2014 & \cite{AU-Whr} \\
AU-Wom & EBF & -37.4222 & 144.0944 & Unk & 2010-2012 & \cite{AU-Wom} \\
BE-Lon & CRO & 50.5516 & 4.7461 & Cfb & 2004-2014 & \cite{BE-Lon} \\
BE-Vie & MF & 50.3051 & 5.9981 & Cfb & 1996-2014 & \cite{BE-Vie} \\
BR-Sa3 & EBF & -3.0180 & -54.9714 & Am & 2000-2004 & \cite{BR-Sa3} \\
CA-Qfo & ENF & 49.6925 & -74.3421 & Dfc & 2003-2010 & \cite{CA-Qfo} \\
CA-SF1 & ENF & 54.4850 & -105.8176 & Dfc & 2003-2006 & \cite{CA-SF1} \\
CA-SF2 & ENF & 54.2539 & -105.8775 & Dfc & 2001-2005 & \cite{CA-SF1} \\
CH-Cha & GRA & 47.2102 & 8.4104 & Unk & 2005-2014 & \cite{CH-Cha} \\
CH-Dav & ENF & 46.8153 & 9.8559 & Unk & 1997-2014 & \cite{CH-Dav} \\
CH-Fru & GRA & 47.1158 & 8.5378 & Unk & 2005-2014 & \cite{CH-Fru} \\
DE-Geb & CRO & 51.1001 & 10.9143 & Unk & 2001-2014 & \cite{DE-Geb} \\
DE-Gri & GRA & 50.9500 & 13.5126 & Cfb & 2004-2014 & \cite{DE-Gri} \\
DE-Hai & DBF & 51.0792 & 10.4530 & Unk & 2000-2012 & \cite{DE-Hai} \\
DE-Kli & CRO & 50.8931 & 13.5224 & Cfb & 2004-2014 & \cite{DE-Gri} \\
DE-Lkb & ENF & 49.0996 & 13.3047 & Unk & 2009-2013 & \cite{DE-Lkb} \\
DE-Obe & ENF & 50.7867 & 13.7213 & Cfb & 2008-2014 & {\textendash} \\
DE-Seh & CRO & 50.8706 & 6.4497 & Unk & 2007-2010 & \cite{DE-Seh} \\
DE-Tha & ENF & 50.9624 & 13.5652 & Cfb & 1996-2014 & \cite{DE-Tha} \\
DK-Sor & DBF & 55.4859 & 11.6446 & Unk & 1996-2014 & \cite{DK-Sor} \\
FI-Hyy & ENF & 61.8474 & 24.2948 & Unk & 1996-2014 & \cite{FI-Hyy} \\
FI-Sod & ENF & 67.3619 & 26.6378 & Unk & 2001-2014 & \cite{FI-Sod} \\
FR-Gri & CRO & 48.8442 & 1.9519 & Cfb & 2004-2013 & \cite{FR-Gri} \\
FR-LBr & ENF & 44.7171 & -0.7693 & Unk & 1996-2008 & \cite{FR-LBr} \\
IT-Col & DBF & 41.8494 & 13.5881 & Unk & 1996-2014 & \cite{IT-Col} \\
IT-Cpz & EBF & 41.7052 & 12.3761 & Unk & 1997-2009 & \cite{IT-Cpz} \\
IT-Lav & ENF & 45.9562 & 11.2813 & Unk & 2003-2014 & \cite{IT-Lav} \\
IT-MBo & GRA & 46.0147 & 11.0458 & Unk & 2003-2013 & \cite{IT-MBo} \\
IT-Noe & CSH & 40.6061 & 8.1515 & Unk & 2004-2014 & \cite{IT-Noe} \\
IT-Ren & ENF & 46.5869 & 11.4337 & Unk & 1998-2013 & \cite{IT-Ren} \\
IT-Ro2 & DBF & 42.3903 & 11.9209 & Unk & 2002-2012 & \cite{IT-Ro2} \\
IT-SRo & ENF & 43.7279 & 10.2844 & Unk & 1999-2012 & \cite{IT-SRo} \\
IT-Tor & GRA & 45.8444 & 7.5781 & Unk & 2008-2014 & \cite{IT-Tor} \\
NL-Loo & ENF & 52.1666 & 5.7436 & Unk & 1996-2013 & \cite{NL-Loo} \\
RU-Fyo & ENF & 56.4615 & 32.9221 & Unk & 1998-2014 & \cite{RU-Fyo} \\
US-AR1 & GRA & 36.4267 & -99.4200 & Dsa & 2009-2012 & \cite{US-AR1} \\
US-AR2 & GRA & 36.6358 & -99.5975 & Dsa & 2009-2012 & \cite{US-AR1} \\
US-ARM & CRO & 36.6058 & -97.4888 & Cfa & 2003-2012 & \cite{US-ARM} \\
US-Blo & ENF & 38.8953 & -120.6328 & Csa & 1997-2007 & \cite{US-Blo} \\
US-KS2 & CSH & 28.6086 & -80.6715 & Cwa & 2003-2006 & \cite{US-KS2} \\
US-MMS & DBF & 39.3232 & -86.4131 & Cfa & 1999-2014 & \cite{US-MMS} \\
US-Me2 & ENF & 44.4523 & -121.5574 & Csb & 2002-2014 & \cite{US-Me2} \\
US-NR1 & ENF & 40.0329 & -105.5464 & Dfc & 1998-2014 & \cite{US-NR1} \\
US-Ne1 & CRO & 41.1651 & -96.4766 & Dfa & 2001-2013 & \cite{US-Ne1} \\
US-Ne2 & CRO & 41.1649 & -96.4701 & Dfa & 2001-2013 & \cite{US-Ne1} \\
US-Ne3 & CRO & 41.1797 & -96.4397 & Dfa & 2001-2013 & \cite{US-Ne1} \\
US-SRG & GRA & 31.7894 & -110.8277 & Bsk & 2008-2014 & \cite{US-SRG} \\
US-SRM & WSA & 31.8214 & -110.8661 & Bsk & 2004-2014 & \cite{US-SRM} \\
US-Syv & MF & 46.2420 & -89.3477 & Dfb & 2001-2014 & \cite{US-Syv} \\
US-Ton & WSA & 38.4316 & -120.9660 & Csa & 2001-2014 & \cite{US-Ton} \\
US-Var & GRA & 38.4133 & -120.9507 & Csa & 2000-2014 & \cite{US-Var} \\
US-WCr & DBF & 45.8059 & -90.0799 & Dfb & 1999-2014 & \cite{US-WCr} \\
US-Wkg & GRA & 31.7365 & -109.9419 & Bsk & 2004-2014 & \cite{US-Wkg} \\
ZA-Kru & SAV & -25.0197 & 31.4969 & Unk & 2000-2010 & \cite{ZA-Kru} \\
ZM-Mon & DBF & -15.4378 & 23.2528 & Unk & 2000-2009 & \cite{ZM-Mon} \\